The hypoxic response extends lifespan through a bioaminergic and peptidergic neural circuit

Version 1

Abstract

Abstract

A coordinated response to stress is crucial for promoting the short- and long-term health of an organism. The perception of stress, frequently through the nervous system, can lead to physiological changes that are fundamental to maintaining homeostasis.

Activating the response to low oxygen, or hypoxia, extends healthspan and lifespan in C. elegans. However, despite some positive impacts, negative effects of the hypoxic response in specific tissues prevent translation of their benefits in mammals. Thus, it is imperative to identify which components of this response promote longevity. Here, we interrogate the cell-nonautonomous hypoxic response signaling pathway. We find that HIF-1-mediated signaling in ADF serotonergic neurons is both necessary and sufficient for lifespan extension. Signaling through the serotonin receptor SER-7 in the GABAergic RIS interneurons is necessary in this process. Our findings also highlight the involvement of additional neural signaling molecules, including the neurotransmitters tyramine and GABA, and the neuropeptide NLP-17, in mediating longevity effects. Finally, we demonstrate that oxygen- and carbon-dioxide-sensing neurons act downstream of HIF-1 in this circuit.

Together, these insights develop a circuit for how the hypoxic response cell-nonautonomously modulates aging and suggests valuable targets for modulating aging in mammals.

Introduction

The global population is aging at an unprecedented rate, posing significant challenges for healthcare systems worldwide1. The increasing burden of age-related diseases such as cardiovascular disease, diabetes, and neurodegeneration2 highlights the pressing need to improve healthspan, defined as the disease-free period of life3. Work over the past several decades has led to a better understanding of the conserved molecular and cellular mechanisms that regulate lifespan. One promising avenue has been the identification of environmental stressors that, when applied in a controlled manner, activate stress-response pathways to promote longevity. These interventions, including calorie restriction, exercise, hypoxia, heat shock, and cold exposure, extend lifespan and healthspan4–7. These stress-response pathways induce hormesis, a phenomenon in which a low dose of a stressor results in adaptive beneficial effects on cellular function8,9.

Hormesis and the underlying stress-responses that promote longevity converge on cellular pathways, including oxidative stress, insulin signaling, autophagy, and protein homeostasis10, to promote longevity.

Multicellular organisms secrete a highly conserved set of neuromodulators including neurotransmitter and neuropeptide signals from the central nervous system. These signals integrate information about the organism’s external environment and internal state, and coordinate a physiological response in peripheral tissues. Within the context of aging, bioamine neurotransmitters contribute to the activation of many well-studied stress response mechanisms that promote longevity. For example, serotonin signaling modulates the mitochondrial unfolded protein response in C. elegans11, and regulates how food perception contributes to dietary restriction-mediated longevity in worms12 and flies13.

Dopamine also modifies the longevity benefits of dietary restriction in invertebrates12,14, and polymorphisms in the dopamine D4 receptor are associated with longevity in humans15. Finally, adrenaline and its invertebrate analog, tyramine, have been linked to changes in the aging rate in both nematodes14,16–18 and mammals19. Hormone signals such as insulin20–22, growth factors23,24, and GnRH14,25 also contribute to longevity pathways across taxa. This growing body of work has led to great interest in manipulating longevity-promoting neural circuits to improve health26. While significant research effort focuses on pathways like autophagy, insulin signaling, and proteostasis in the context of aging, the role of hypoxia remains relatively underexplored. Currently, no longevity-promoting agents have been reported to target the hypoxic response.

The hypoxic response is highly conserved across species27–29, highlighting the relevance of this pathway in other organisms, including humans. In C. elegans, activating the hypoxic response by knocking down or mutating the E3 ubiquitin ligase von hippel-lindau-1 (VHL-1)30 protein or stabilizing the hypoxia-inducible factor-1 (HIF-1) transcription factor6,31 is sufficient to extend lifespan. All of these interventions—environmental hypoxia, VHL-1 knockdown, and HIF-1 stabilization—require the intestinal enzyme flavin-containing monooxygenase-2 (fmo-2) to extend lifespan in C. elegans31. Hypoxic conditions also extend lifespan in a short-lived progeria mouse model32, and epidemiological studies indicate a correlation between hypoxia exposure and longevity33,34. However, the physiological changes induced by the hypoxic response are broad and involve adaptations such as increased vascularization, metabolic rewiring, and changes in cell survival pathways. In mammals, some of these same adaptations can be detrimental, as mutations in components of the hypoxic response have been linked to conditions like cancer and cardiovascular disease35–38. This presents a key challenge: while hypoxia-induced longevity in a post-mitotic invertebrate model like C. elegans is promising, some of the mechanisms involved may not be directly translatable to mammals without causing deleterious side effects. Therefore, a mechanistic understanding of the individual cells, neural circuits, and pathways that mediate the beneficial but not detrimental effects of hypoxia on aging is essential to determine whether this circuit could be leveraged to improve human health.

Our previous work in C. elegans identified that stabilization of HIF-1 in neurons is sufficient to extend lifespan through the serotonin receptor, SER-7. This pathway eventually leads to the induction of fmo-2, a longevity gene expressed in the intestine31,39. In this study, we uncover key neural components of the hypoxic response longevity circuit. Within this circuit, we identify individual cells, signals, and receptors necessary and/or sufficient to extend lifespan downstream of the hypoxic response. More specifically, we find serotonin signaling in the ADF serotonergic neurons is both necessary and sufficient to extend lifespan through the hypoxic response. This pathway signals through the serotonin receptor SER-7 in the RIS interneuron, which is also essential for hypoxia-mediated longevity. We further demonstrate additional neurotransmitters (GABA and tyramine), and a neuropeptide (NLP-17) are critical for mediating these longevity effects. Finally, we identify that oxygen sensing neurons (URX, AQR, PQR and BAG) act downstream of neuronal HIF-1 in this circuit. Our insights into this longevity pathway provide a mechanistic understanding of how the hypoxic response delays aging and improves health.

Results

Serotonin signaling through the ADF neuron and the SER-7 receptor are necessary and sufficient for the hypoxic response to extend lifespan

Induction of the hypoxic response by targeted genetic manipulations can increase lifespan in C. elegans. These manipulations include decreasing activity of the VHL-1 E3 ubiquitin ligase that targets the transcription factor HIF-1 for degradation under hypoxia30 or by a mutation that stabilizes HIF-1 (HIF-1P621A)6,31 (Fig. 1A). We previously found that stabilizing HIF-1 in serotonergic neurons is sufficient to extend lifespan31. To determine which serotonergic neurons initiate HIF-1-mediated longevity, we generated strains with a nondegradable HIF-1 variant (HIF-1P621A)31,40 expressed under promoters specific to each of C. elegans’ three primary serotonergic neuron types—the ADF, NSM, and HSN neurons41. These transgenic worms were then crossed into the hif-1 null background to ensure that HIF stabilization in one neuron type could not feedback to modify HIF-1 activity in other cells. When we measured the lifespan of each strain relative to WT and hif-1 knockout controls, we observed that stabilizing HIF-1 in the ADF or NSM neurons significantly extended lifespan by 26% and 23%, respectively (Fig. 1B-C). HIF-1 stabilization in the HSN neurons had a smaller effect but significantly extended lifespan by 9% (Fig. 1D). This result indicates that modifying signaling in any serotonergic neuron is sufficient to induce some level of hypoxic response-mediated longevity. These data suggest that serotonergic neurons can partially substitute for each other to promote longevity in this pathway. In summary, the ADF and/or NSM serotonergic neurons play an essential role in hypoxic response-mediated longevity.

ADF serotonergic neurons are necessary and sufficient to extend lifespan downstream of the hypoxic response.

(A) Diagram of the conserved hypoxic response and genetic approaches to activate it, vhl-1 knockdown and HIF-1 stabilization. (B-D) Survival curves of WT, hif-1(ia4) knockout, hif-1(ia4); ADF::HIF-1S (B), hif-1(ia4); NSM::HIF-1S (C), and hif-1(ia4); HSN::HIF-1S (D) worms. N ≥ 256 (B), N ≥ 188 (C), and N ≥ 188 (D) worms per condition. (E) Survival curves of WT and ADF:tph-1 knockout worms on empty vector (EV) or vhl-1 RNAi. N ≥ 143 worms per condition. (F) Survival curves of WT and NSM:tph-1 knockout worms on empty vector (EV) or vhl-1 RNAi. N ≥ 193 worms per condition. (G) Survival curves of WT and tph-1(mg280); ADF:tph-1 rescue worms on empty vector (EV) or vhl-1 RNAi. N ≥ 178 worms per condition. (H) Survival curves of WT and tph-1(mg280); NSM:tph-1 rescue worms on empty vector (EV) or vhl-1 RNAi. N ≥ 160 worms per condition. (I) Gene expression of fmo-2 in WT and hif-1(ia4); ADF:HIF-1S worms. Significance in panels B-H is from a log-rank test comparing median survival. Significance in panel I is from a Student’s t-test (unpaired, two-tailed). In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001.

To further test the necessity of the most impactful (for lifespan) serotonergic neurons in hypoxia-mediated longevity, we used transgenic animals in which tph-1, the rate-limiting enzyme in serotonin synthesis, was knocked out exclusively in the ADF or the NSM serotonergic neurons12,42. We then measured the lifespans of these strains on empty vector (EV) and vhl-1 RNAi to determine the necessity of ADF and NSM serotonergic signaling in vhl-1-mediated longevity. Successful RNAi knockdown was confirmed with qPCR validation (Fig. S1A). We found that serotonin synthesis in ADF but not NSM neurons was required for vhl-1 knockdown to extend lifespan (Fig. 1E-F). This indicates that while HIF stabilization in multiple serotonergic neurons can extend lifespan, ADF serotonin signaling is required for lifespan extension in response to vhl-1 knockdown.

tph-1 is required for vhl-1 mediated longevity.

(A) Gene expression of vhl-1 in worms on vhl-1 RNAi compared to control worms on empty vector (EV) RNAi for two generations. N ≥ 200 worms per replicate, or 1200 worms per condition. The top of the bar represents the mean of the population and error bars indicate standard error of the mean (SEM). (B) Survival curve of WT, vhl-1(ok161), tph-1(mg280), and vhl-1(ok161); tph-1(mg280) worms. N ≥ 338 worms per condition. Significance is from a log-rank test comparing median survival. In all panels, NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. Three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

To test if serotonin production in a single neuron type is sufficient for the hypoxic response to extend lifespan, we rescued tph-1 expression in a tph-1 null background using promoters specific to the ADF or the NSM neurons12. The lifespans of these strains on EV or vhl-1 RNAi were then measured. As expected and consistent with our previous publication31, the lifespan of the tph-1 knockout strain was not extended by vhl-1 knockdown (Fig. S1B). Consistent with the ADF neuron’s requirement for hypoxic response-mediated longevity (Fig. 1E-F), serotonin synthesis in the ADF neurons was sufficient for vhl-1 knockdown to extend lifespan while NSM serotonin synthesis was not (Fig. 1G-H). The enzyme fmo-2 is induced in the intestine under hypoxia and is required for lifespan extension by environmental hypoxia, HIF-1 stabilization, or vhl-1 depletion31. To test whether ADF-specific HIF-1 stabilization also induces fmo-2, we measured fmo-2 transcription in this strain using qPCR. We found that ADF HIF-1 stabilization significantly induced fmo-2 expression relative to WT controls (Fig. 1I), suggesting a model in which ADF signaling extends lifespan through the same mechanism as vhl-1 knockdown and environmental hypoxia. Together, these data indicate that while HIF-1 stabilization in any serotonergic neuron is sufficient to extend lifespan, only ADF serotonin production is necessary and sufficient for vhl-1-mediated longevity. This suggests that under normal physiological conditions the serotonergic ADF neurons propagate a signal in response to whole body stabilized HIF-1.

After identifying ADF neurons as the most central serotonergic neurons in hypoxic response-mediated longevity, we next sought to determine the downstream receptor responding to serotonin signaling from ADF neurons. C. elegans have six known serotonin receptors—SER-1, SER-4, SER-5, SER-7, MOD-1, and LGC-50. Of these six receptors, our previous findings showed that ser-7 expression is required for hypoxia to induce fmo-2 and extend lifespan31. SER-7 is a G protein-coupled receptor (GPCR) with high sequence identity to the 5-HT7 receptor in mammals43,44. SER-7 is thought to be expressed in 27 of C. elegans’ 302 neurons45 and contributes to many aspects of physiology such as reproduction and pharyngeal pumping43. Consequently, broad manipulation of SER-7 signaling is not an ideal method for exclusively modifying aging. To identify more specific targets for lifespan extension within this pathway, we asked whether ser-7 expression in any subset of neurons is sufficient to rescue the hypoxic response in a ser-7 null background.

We created transgenic strains with ser-7 expression under the control of 11 different cell-specific promoters in a ser-7 null background. These promoters were selected to rescue ser-7 expression in the following neuronal populations: whole-body rescue (ser-7p::ser-7), interneuron rescue (glr-1p::ser-7), bioaminergic neuron rescue (cat-1p::ser-7), glutamatergic neuron rescue (eat-4p::ser-7), GABAergic neuron rescue (unc-47p::ser-7), cholinergic neuron rescue (unc-17p::ser-7), sensory neuron rescue (osm-6p::ser-7), GABAergic motor neuron rescue (unc-25p::ser-7), cholinergic motor neuron rescue (acr-2p::ser-7), M3 and M4 neuron rescue (ceh-28p::ser-7), and intestinal rescue (vha-6p::ser-7) in a ser-7 null background. To measure whether ser-7 expression restores a WT-like hypoxic response, we injected each construct into ser-7 null worms crossed with a single-copy transcriptional reporter for fmo-2 (fmo-2p::mCherry). Because fmo-2 is induced by and required for hypoxia-mediated longevity31, we utilized its induction as an efficient screening tool to identify components of this longevity pathway.

We first confirmed that ser-7 knockout attenuates vhl-1-mediated fmo-2 induction (Fig. 2A). The whole-body rescue of SER-7 expression under the endogenous ser-7 promoter restored fmo-2 induction to 91% of WT controls (Fig. 2A). Of the 10 cell-specific ser-7 rescues, expression in interneurons, bioaminergic neurons, glutamatergic neurons, or GABAergic neurons fully rescued vhl-1-mediated fmo-2 induction to ≥ 100% of the WT control response (Fig. 2A, pink bars). ser-7 rescue in GABAergic motor neurons or in sensory neurons showed a partial rescue (defined as a 50-99% increase from fmo-2 induction in the ser-7 knockout) (Fig. 2A, orange bars). In contrast, rescuing ser-7 in cholinergic motor neurons, cholinergic neurons, M3 and M4 neurons, or in the intestine did not rescue (Fig. 2A, gray bars). Taken together, 6 of the 10 cell-specific ser-7 rescues at least partially restored vhl-1-mediated fmo-2 induction, including some that have no known overlap in ser-7 expression patterns. Collectively, these data suggest that serotonin receptor ser-7 expression in multiple neurons may be sufficient to convey the hypoxic signal to the intestine and induce fmo-2.

ser-7 expression in the GABAergic RIS neuron is required for hypoxia to extend lifespan.

(A) Quantification of fmo-2p::mCherry (WT positive control), fmo-2p::mCherry; ser-7(tm1325) knockout (negative control), and fmo-2p::mCherry; ser-7(tm1325) knockouts with ser-7 rescued in the whole body (ser-7p::ser-7), interneurons (glr-1p::ser-7), bioaminergic neurons (cat-1p::ser-7), glutamatergic neurons (eat-4p::ser-7), GABAergic neurons (unc-47p::ser-7), GABAergic motor neurons (unc-25p::ser-7), sensory neurons (osm-6p::ser-7), cholinergic motor neurons (acr-2p::ser-7), cholinergic neurons (unc-17p::ser-7), M3 & M4 neurons (ceh-28p::ser-7), and the intestine (vha-6p::ser-7) on empty vector (EV, WT control shown) or vhl-1 RNAi (all strains shown). Bar height indicates the mean fluorescence on vhl-1 RNAi of each genotype normalized to the empty vector control value from that genotype. The dashed line indicates the vhl-1-mediated fmo-2 induction of WT positive control. N ≥ 45 worms per condition. Error bars indicate SEM. Significance is from a two-way ANOVA (fmo-2 induction ∼ genotype*RNAi) and post-hoc Tukey HSD test (unpaired, two-tailed). Stars signify results from this post-hoc test comparing the fmo-2 induction of each strain on EV RNAi the fmo-2 induction of that strain on vhl-1 RNAi. (B) A table showing the 10 ser-7 expressing candidate neurons and their expression patterns among the ser-7 cell-specific strains that successfully rescued fmo-2 induction in Fig. 2A. Pink rows indicate rescues that full restored vhl-1 mediated fmo-2 induction, orange rows indicate rescues that partially restored vhl-1-mediated fmo-2 induction, and gray rows signify unsuccessful rescue constructs from the data in Fig. 2A. (C) Survival curve of WT and RIS ablation (Ex[srsx-18p::caspase-3(p12)-nz]; Ex[srsx-18p::cz-caspase-3 (p17)]; srsx-18p::GFP) worms on empty vector (EV) and vhl-1 RNAi. N ≥ 192 worms per condition. (D) Survival curve of WT and ser-7 (tm1325); ser-7 rescue in the RIS neuron (flp-11p::ser-7) worms on EV and vhl-1 RNAi. N ≥ 144 worms per condition. Significance in panels C-D is from a log-rank test comparing median survival. In all panels, NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

We compiled a table containing the top candidates for the ser-7 neuron based on which rescue constructs restored ser-7 expression (simplified table in Fig. 2B, full table in Fig. S2A). We hypothesized that neurons present in many of the successful rescue constructs would be most important in the hypoxia longevity circuit. Conversely, we hypothesized that candidate neurons within rescues that did not restore a WT-like hypoxia response are less likely to be important (Fig. 2B; Fig. S2A). We acknowledge that this analysis strategy is biased toward neurons rescued in a greater proportion of the transgenic strains, which could lead to false negatives. From this analysis, we identified 10 neurons that were rescued in 3-4 of the 6 successful rescue constructs (Fig. 2B). Five of these neurons (DVB, FLP, M1, RME_LR, and RME_DV) were also present in unsuccessful rescue constructs, so they were not considered for follow-up experiments. Of the remaining five neurons (RIS, AUA, AVC, LUA, and VD_DD), the top candidate was the RIS neuron, because RIS was present in the highest number (4 as shown in Fig. 2B) of successful rescue constructs.

ser-7 expression in each ser-7 expressing neuron is rescued by at least one construct, and ser-7 is required for vhl-1-mediated longevity.

(A) A table showing all ser-7 expressing candidate neurons, and which constructs in Fig. 2A restored ser-7 expression in these neurons. Pink rows indicate rescues that fully restored vhl-1 mediated fmo-2 induction, orange rows indicate rescues that partially restored vhl-1-mediated fmo-2 induction, and gray rows signify unsuccessful rescue constructs from the data in panel A. (B) Survival curve of WT, and ser-7(tm1325) worms on empty vector (EV) and vhl-1 RNAi. N ≥ 259 worms per condition. Significance is from a log-rank test comparing median survival. Cox Regression for an interaction between the effect of ser-7 (tm1325) knockout and vhl-1 RNAi knockdown on lifespan. p = 0.003, **. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. Three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

To validate whether the RIS neuron is a key signaling cell in hypoxia-mediated longevity, we first genetically ablated the RIS neuron to test its necessity. Our results showed that vhl-1 RNAi does not extend lifespan of RIS ablated worms, indicating that the RIS neuron is completely required for vhl-1 knockdown to extend lifespan (Fig. 2C). To test the sufficiency of RIS signaling, we next rescued ser-7 expression in the RIS neuron in the ser-7 null background under the flp-11 promoter. flp-11 is expressed in the RIS neuron46 and uv1 neuroendocrine cells47 that do not endogenously express ser-745. We first confirmed that ser-7 knockout worms are partially required for vhl-1 RNAi to extend lifespan (Fig. S2B, Cox Regression, p = 0.003, **). Significantly, ser-7 rescue in the RIS neuron (ser-7; flp-11p::ser-7) restores the lifespan extension by vhl-1 RNAi to the same degree as WT worms (Fig. 2D). These data suggest that ser-7 expression in the RIS neuron only is sufficient for vhl-1-mediated longevity. However, one limitation of this approach is that non-physiological expression of ser-7 in the uv1 cells could play a role in restoration of a WT-like hypoxic response. Together, these data are consistent with a model where hypoxic conditions stabilize HIF-1 in the ADF neurons, which modify serotonin signaling to the SER-7 receptor on the RIS interneuron.

Tyraminergic signaling from the RIM neuron and the tyramine receptor TYRA-3 are required for vhl-1 knockdown to extend lifespan

After identifying neuron subtypes involved in serotonin signaling downstream of the hypoxic response, we next set out to identify whether other neurotransmitters act in the pathway. To answer this question, we obtained mutants deficient in the production of each type of neurotransmitter in C. elegans (Fig. 3A) and measured their lifespan on control and vhl-1 RNAi. Out of the six neurotransmitters we tested, blocking the production of GABA (unc-25, Fig. 3B) or tyramine + octopamine (tdc-1, Fig. 3C) attenuated longevity on vhl-1 RNAi. The requirement of GABA is consistent with our finding that ser-7 expression on the GABAergic RIS neuron is also necessary for hypoxic response-mediated longevity (Fig. 2), connecting the serotonergic and GABAergic components of this circuit.

GABA and tyramine synthesis are required for the hypoxic response to extend lifespan.

(A) Summary table of the necessity of each C. elegans neurotransmitter for vhl-1 RNAi-mediated longevity. (B-E) Survival curves of WT and unc-25(e156) (GABA deficiency) (B), tdc-1(n3419) (tyramine + octopamine deficiency) (C), tbh-1(n3247) (octopamine deficiency) (D), and RIC ablation (Ex[tbh-1p::caspase-3(p12)-nz]; Ex[tbh-1p::cz-caspase-3 (p17)]; tbh-1p::GFP) (E) worms on empty vector (EV) and vhl-1 RNAi. N ≥ 310 (B), ≥ 201 (C), ≥ 258 (D), and ≥ 258 (E) worms per condition. Significance in panels B-E is from a log-rank test comparing median survival. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

Tyramine (analogous to mammalian adrenaline48) is synthesized from tyrosine by tyrosine decarboxylase (tdc-1) and can be further converted to octopamine (analogous to mammalian noradrenaline) by tyramine β-hydroxylase (tbh-1) in C. elegans49. Therefore, the tdc-1 knockout is deficient in both tyramine and octopamine synthesis (Fig. 3C) and the tbh-1 knockout (Fig. 3D) is only deficient in octopamine synthesis. Since worms lacking only octopamine (tbh-1, Fig. 3D) were long-lived on vhl-1 RNAi but worms lacking both tyramine and octopamine (tdc-1, Fig. 3C) were not, we can conclude that tyramine is required for the hypoxic response to extend lifespan. To further test whether tyramine but not octopamine is involved in this pathway, we generated a strain in which the primary octopamine-producing neuron, RIC49, is genetically ablated. We found that the RIC neuron is not required for vhl-1-mediated longevity (Fig. 3E), further supporting that tyramine but not octopamine is a key neurotransmitter in the hypoxic response longevity circuit.

In addition to octopamine (tbh-1, Fig. S3C), we also observed that acetylcholine (unc-17, Fig. S3A) and glutamate (eat-4, Fig. S3B) were not required for lifespan extension by vhl-1 RNAi. Our lifespans of the dopamine synthesis mutant cat-2 indicated that dopamine is inconsistently required for the hypoxic response to extend lifespan. In three biological replicates, we observed a full requirement, partial requirement, and no requirement for dopamine synthesis in vhl-1 RNAi-mediated longevity. When analyzed together, the net result suggests that dopamine is partially required for vhl-1-mediated longevity (Fig. S3C, Cox Regression for interaction between genotype and vhl-1 RNAi, p < 0.0001, ****). We did not follow up on dopamine due to the inconsistency. Together, the results from this neurotransmitter screen support the conclusion that in addition to serotonin, both GABA and tyramine play a role in hypoxic response-mediated longevity.

Acetylcholine, glutamate, and dopamine synthesis are not fully required for vhl-1 RNAi to extend lifespan.

(A-C) Survival curves of WT and unc-17(e113) (acetylcholine deficiency) (A), eat-4(ky5) (glutamate deficiency) (B), and cat-2(n4547) (dopamine deficiency) (C) worms on empty vector (EV) and vhl-1 RNAi. N ≥ 213 (A), ≥ 274 (B), and ≥ 270 (C) worms per condition. (C) Cox Regression for an interaction between genotype and vhl-1 RNAi, p < 0.0001, ****. Significance in all panels is from a log-rank test comparing median survival. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

Nematodes synthesize GABA in 26 of 302 neurons while tyramine is only synthesized in two cells50. As a result, identifying highly specific components of tyraminergic signaling is more feasible than mapping GABAergic signaling components. Therefore, we sought to determine the tyramine-producing neuron(s) and tyramine receptor(s) that contribute to hypoxic response-mediated longevity. In C. elegans, the two canonically tyraminergic cell types (the RIM neuron and the uv1 neuroendocrine cells) express tdc-1 but not tbh-150. To determine which tyraminergic cell(s) act in the hypoxic response pathway, we expressed tdc-1 under promoters specific to either RIM or uv1 in a tdc-1 null background14. We then measured whether the lifespans of these tyramine rescue strains could be extended by vhl-1 RNAi. Our results showed that vhl-1 RNAi extended lifespan in the RIM rescue strain (Fig. 4A), but not in the uv1 rescue (Fig. 4B). This finding demonstrates that tyramine synthesis in the RIM neuron is sufficient for vhl-1 RNAi to extend lifespan. Notably, both tyramine rescue constructs lived longer than WT controls (Fig. 4A-B). This could indicate that altering tyramine signaling modifies lifespan both within and independently of the vhl-1-mediated longevity pathway.

The RIM neuron and the tyramine receptor tyra-3 act in the hypoxic response-mediated longevity pathway.

(A-B) Survival curves of WT, RIM rescue (tdc-1 (n3419); Ex[ocr-4p::tdc-1]) (A), and uv1 tdc-1 rescue (tdc-1 (n3419); Ex[gcy-13p::tdc-1]) (B) strains on empty vector (EV) and vhl-1 RNAi. N ≥ 126 (A) and N ≥ 140 (B) worms per condition. (C) Survival curves of TU3311(unc-119p::sid-1) and vhl-1(ok161); TU3311(unc-119p::sid-1) worms on empty vector (EV) and tyra-3 RNAi. N ≥ 215 worms per condition. Significance in all panels is from a log-rank test comparing median survival. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

To identify the tyramine receptor(s) necessary in the hypoxic response-mediated longevity pathway, we used RNAi to knockdown each of C. elegans’ four known tyramine receptors51–54 (Fig. 4C). Since tyramine receptors are mostly expressed in neurons, and RNAi uptake into some neurons is less efficient than other cell types, we used a strain with enhanced neuronal RNAi uptake (TU3311, unc-119p::sid-1) crossed into vhl-1 knockout worms. This neural enhanced RNAi uptake strain was not used previously in this work for vhl-1 RNAi experiments because vhl-1 RNAi knockdown extends lifespan in WT worms31.

The generated vhl-1; unc-119p::sid-1 strain was long-lived compared to the control strain unc-119p::sid-1 (Fig. 4C). Successful RNAi knockdown was validated via qPCR (Fig. S4A). Our lifespan results show that the tyramine receptor tyra-3 was fully required for vhl-1- mediated longevity (Fig. 4C), but the tyramine receptors tyra-2 and ser-2 were not (Fig. S4B-C). Another tyramine receptor, Igc-55, showed a partial requirement for vhl-1-mediated longevity (Fig. S4D). Although there was a lifespan extension by vhl-1 on Igc-55 RNAi (Fig. S4D), we did observe a significant interaction between lgc-55 knockdown and genotype on lifespan (Cox Regression, p < 0.0001, ****), indicating a partial requirement. In summary, these data indicate that tyramine signaling from the RIM neuron to the TYRA-3 receptor is required for hypoxic response-mediated longevity, while the LGC-55 receptor may play a partial role. Interestingly, TYRA-3 is highly expressed on the low oxygen/high carbon dioxide-sensing BAG neuron as well as in the intestine45,55, the site of fmo-2 induction during hypoxia31.

tyra-2 and ser-2 are not necessary for vhl-1-mediated longevity, while lgc-55 is partially required.

(A) Gene expression of tyra-3, tyra-2, ser-2, and lgc-55 in TU3311(unc-119p::sid-1) worms raised on RNAi targeting each gene compared to control worms raised on empty vector (EV) RNAi for two generations. N ≥ 300 worms per replicate, or 900 worms per condition. The top of the bar represents the mean of the population and error bars indicate standard error of the mean (SEM). (B-D) Survival curves of TU3311(unc-119p::sid-1) and vhl-1(ok161); TU3311(unc-119p::sid-1) worms on empty vector (EV) and tyra-2 (B), ser-2 (C), or lgc-55 (D) RNAi. Cox Regression for an interaction between lgc-55 knockdown and vhl-1 knockout on lifespan in the TU3311background strain. p < 0.0001, ****. N ≥ 215 worms per condition. Significance is from a log-rank test comparing median survival. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

Oxygen and carbon dioxide sensing neurons act downstream of serotonin signaling

Given that tyra-3 is highly expressed in the canonical low-oxygen sensing neuron BAG45, we next explored whether neurons responsible for sensing high- and low-oxygen conditions are essential for hypoxic response-mediated longevity. C. elegans have four oxygen sensory neurons: the URX, PQR, and AQR neurons detect high levels of oxygen55, the BAG neuron responds to low levels of oxygen56 and high levels of carbon dioxide57 (Fig. 5A). We hypothesized that these oxygen sensing cells may perceive hypoxic conditions and help initiate the hypoxic response. Alternatively, these oxygen sensing neurons could modify behavior in response to hypoxia but not play a role in longevity. We found that vhl-1 knockdown was unable to extend lifespan when we genetically ablated the three high-oxygen sensing neurons (URX, AQR, PQR, Fig. 5B). Interestingly, the low O2/high CO2 responsive BAG neurons were also required for vhl-1-mediated longevity (Fig. 5C). We next crossed the URX/AQR/PQR ablation worms or BAG ablation worms into C. elegans with HIF-1 stabilized in the ADF neurons. Ablation of either the URX/PQR/AQR or the BAG neurons abrogated lifespan extension in the ADF HIF-1 stabilized worms. This suggests that oxygen sensing neurons act downstream of ADF HIF-1 stabilization in this pathway (Fig. 5D-E).

High- and low-oxygen sensing neurons are important for hypoxic response-mediated longevity.

(A) Diagram of oxygen sensing neurons in C. elegans. (B) Survival curves of WT and URX/PQR/AQR ablated worms (qaIs2241 [gcy-36::egl-1 + gcy-35::GFP + lin-15(+)]) on empty vector (EV) and vhl-1 RNAi. N ≥ 279 worms per condition. (C) Survival curves of WT and BAG ablated worms (Ex[gcy-31p::caspase-3(p12)-nz]; Ex[gcy-31p::cz-caspase-3 (p17)]; gcy-31p::GFP) on empty vector (EV) and vhl-1 RNAi. N ≥ 188 worms per condition. (D) Survival curves of WT, hif-1(ia4); ADF-HIF-1S, URX/PQR/AQR (qaIs2241 [gcy-36::egl-1 + gcy-35::GFP + lin-15(+)]) ablation, and URX/PQR/AQR (qaIs2241 [gcy-36::egl-1 + gcy-35::GFP + lin-15(+)]) ablation; hif-1(ia4); ADF-HIF-1S worms. N ≥ 150 worms per condition. (E) Survival curves of WT, hif-1(ia4); ADF-HIF-1S, BAG (Ex[gcy-31p::caspase-3(p12)-nz]; Ex[gcy-31p::cz-caspase-3 (p17)]; gcy-31p::GFP) ablation, and BAG (Ex[gcy-31p::caspase-3(p12)-nz]; Ex[gcy-31p::cz-caspase-3 (p17)]; gcy-31p::GFP) ablation; hif-1(ia4); ADF-HIF-1S worms. N ≥ 186 worms per condition. Significance in all panels is from a log-rank test comparing median survival. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

The neuropeptide NLP-17 and its receptors contribute to vhl-1-mediated longevity

After identifying neurons, neurotransmitters, and neuroreceptors acting in the hypoxic response longevity circuit, we wondered how this neural circuit ultimately propagates information about hypoxic conditions to the intestine and induces the pro-longevity gene fmo-2. The C. elegans nervous system does not directly innervate the intestine. Instead, neurosignaling molecules bind to receptors on peripheral tissues through packaging in dense-core vesicles and subsequent release into and diffusion through pseudocoelomic fluid58,59. These dense-core vesicles are packaged with a wide variety of neuropeptide signals, although there is also evidence that bioaminergic neurotransmitters like serotonin, dopamine, and adrenaline/noradrenaline can also signal through this dense-core vesicle mechanism59–61.

We first tested if inducing the hypoxic response through vhl-1 RNAi extends lifespan in a mutant strain lacking the dense-core vesicle packaging gene unc-31. We observed that unc-31 was required for vhl-1 knockdown to extend lifespan (Fig. 6A). To examine whether neuropeptide signals carried by dense-core vesicles are involved, we knocked down two neuropeptide processing enzymes—egl-3 and egl-2162. Successful knockdown was validated with qPCR (Fig. S5A). Interestingly, egl-21 but not egl-3 was required for vhl-1- mediated longevity (Fig. 6B-C). This could be because egl-21 is the only enzyme known to perform the carboxypeptidase E cleavage step of neuropeptide processing63, while there are four known proprotein convertases with similar function to egl-362,64.

The neuropeptide nlp-17 and its receptors npr-37 and npr-43 are required for hypoxic response-mediated longevity.

(A) Survival curves of unc-31(e169) and WT worms on empty vector (EV) and vhl-1 RNAi. N ≥ 242 worms per condition. (B-C) Survival curves of TU3311 (unc-119p::sid-1) and vhl-1(ok161); TU3311 (unc-119p::sid-1) worms on empty vector (EV) and egl-3 (B) or egl-21 (C) RNAi. N ≥ 209 worms (B) and N ≥ 185 worms (C) per condition. (D) Summary table of two sets of neuropeptide ligands/receptors that blunted vhl-1- mediated lifespan. (E-G) Survival curves of TU3311 (unc-119p::sid-1) and vhl-1(ok161); TU3311 (unc-119p::sid-1) worms on empty vector (EV) and npr-37 (E), npr-43 (F), or nlp-17 (G) RNAi. N ≥ 288 worms (E), N ≥ 309 worms (F), and N ≥ 309 worms (G) per condition. Significance in panels A-C and E-G are from log-rank test comparing median survival. Cox Regression for a significant interaction between npr-43 knockdown and vhl-1 knockout on lifespan, p <0.01, **. Cox regression for a significant interaction between nlp-17 knockdown and vhl-1 knockout on lifespan, p < 0.05, *. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

We next tested a group of 6 genes corresponding to neuropeptide/receptor binding pairs that blunted hypoxic response-mediated fmo-2 induction (Fig. 6D). These genes, along with their corresponding receptors, were targeted for RNAi and tested for lifespan in TU3311 (unc-119p::sid-1) and TU3311 (unc-119p::sid-1); vhl-1 (ok161) strains. We found that the npr-3 receptor and it’s ligands, flp-15 and flp-21 were not required for vhl-1-mediated longevity. (Fig. S5B-D). However, knocking down the neuropeptide nlp-17 partially blocked vhl-1-mediated longevity (Fig. 6E, Cox regression p < 0.05). Consistently, knocking down nlp-17’s receptors also completely (npr-37, Fig. 6F) or partially (npr-43, Fig. 6G, Cox regression p < 0.001) blocked vhl-1-mediated longevity. Successful knockdown of npr-37 and npr-43 was validated with qPCR (Fig. S5A). The nlp-17 ligand result was validated using a genetic knockout (nlp-17 (ok3461)), which completely prevented vhl-1 knockdown from extending lifespan (Fig. S5E). Together, these data suggest a role for NLP-17 signaling in hypoxic response-mediated longevity.

qPCR validation of RNAi hits from Figure 5, and neuropeptide screen hits that did not validate.

(A-B) Gene expression of (A) egl-3, egl-21, npr-37, and npr-43 in TU3311 (unc-119p::sid-1) worms raised on RNAi targeting each gene compared to control worms raised on empty vector (EV) RNAi for two generations. N ≥ 300 worms per replicate, or 900 worms per condition. The top of the bar represents the mean and error bars represent SEM. (B-D) Survival curves of TU3311 (unc-119p::sid-1) and vhl-1(ok161); TU3311 (unc-119p::sid-1) worms on empty vector (EV) and npr-3 (B), flp-15 (C), or flp-21 (D) RNAi. N ≥ 191 worms (B), N ≥ 181 worms (C), and N ≥ 183 worms (D) per condition. (E) Survival curves of nlp-17(ok3461) on empty vector (EV) or vhl-1 RNAi. N ≥ 275 worms per condition. Significance in panels B-E are from log-rank test comparing median survival. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, 2-3 replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

Taken together, these results support a model in which HIF-1 stabilization in the ADF serotonergic neurons leads to modified serotonin signaling from the ADF to the SER-7 expressing and GABA-producing RIS neuron (Fig. 7). Oxygen sensing neurons act downstream of HIF-1 stabilization in the ADF neuron in the hypoxic response longevity circuit. Tyramine produced by the RIM neuron and neuropeptide (NLP-17) signaling are also required for the hypoxic response to extend lifespan, although it remains unclear whether these signals act upstream, downstream, or in parallel to the serotonergic and GABAergic components in this pathway.

Working model of the hypoxia-mediated longevity pathway.

Discussion

In this study, we interrogated a complex signaling pathway through which genetic induction of the hypoxic response extends lifespan in C. elegans. Within this pathway, we demonstrate that: 1) stabilization of HIF-1 in the ADF serotonergic neurons modifies signaling to the SER-7 receptor on the GABAergic and neuropeptidergic RIS neuron (Fig. 1-2); 2) tyraminergic signaling through the RIM neuron and the TYRA-3 receptor (Fig. 3-4) are required for hypoxic response-mediated longevity; 3) oxygen and carbon-dioxide sensory neurons act downstream of HIF-1 stabilization and serotonergic ADF neurons to extend lifespan (Fig. 5); and 4) in addition to neurotransmitter signals, the neuropeptide NLP-17 and its receptors NPR-37 and NPR-43 are required for the hypoxic response to extend lifespan (Fig. 6). Ultimately, these neural signals converge on the intestine, to activate FMO-2 and extend lifespan (working model summarized in Fig. 7).

The hypoxic response is highly conserved across species, highlighting the potential relevance of this pathway in other organisms, including humans. In nematodes, activating the hypoxic response promotes health and longevity. However, the physiological changes induced by the hypoxic response are broad and involve adaptations that can drive tumorigenesis in mammals. Therefore, a mechanistic understanding of the individual cells, signals, and circuits that mediate the beneficial effects of hypoxia is essential. In this work, the discovery that HIF-1 stabilization in the ADF neurons extends lifespan by 26% demonstrates the potential for targeted manipulations within this pathway to have large effects on lifespan. Additionally, our data identify a previously unknown role for tyramine/adrenaline in the hypoxic response as well as establish NLP-17, a neuropeptide with no previous known function, as a signal in vhl-1-mediated longevity. This work also demonstrates that the oxygen sensing AQR, PQR, URX, and BAG neurons act downstream of the serotonergic ADF neuron in the hypoxic response, indicating an interaction between internal and external oxygen sensing mechanisms in longevity. Together, the mapping of these cells, signaling molecules, and receptors in the hypoxic response lays the foundation to discover druggable targets that selectively modulate aging in humans without negative side effects.

Other studies in C. elegans have also identified neural networks involving the hypoxic response. For example, previous work found a role for HIF-1 in hypoxia-mediated behaviors, such as food-dependent hyperoxia (excessive oxygen) avoidance65. Hyperoxia avoidance requires hif-1 expression both in the neurons and the tyraminergic uv1 neuroendocrine cells. Interestingly, serotonin signaling from the ADF neuron is also required for WT-like hyperoxia avoidance65. While this HIF-1-mediated behavioral circuit has parallels with the HIF-1-mediated longevity circuit interrogated in this work, these pathways do diverge: the key tyraminergic cell type implicated in hyperoxia avoidance is the uv1 gonadal neuroendocrine cells, while the RIM tyraminergic neurons played a role in longevity (Fig. 4). Together, the partial overlap between this behavioral and longevity circuit implies that modifying neural signaling to induce the hypoxic response and extend lifespan may also alter behavioral phenotypes. Future research in this area should interrogate exactly where these and any other HIF-1-mediated circuits overlap and diverge.

While this study identifies many neural signals required for vhl-1 knockdown or knockout to extend lifespan, one key limitation of this work is the potential differences between genetic and environmental methods of inducing the hypoxic response. While vhl-1 knockdown or knockout leads to HIF-1 stabilization by blocking its proteasomal degradation, it also results in hydroxylated but stable HIF-1. This contrasts with environmental hypoxia, in which HIF-1 remains stable because it cannot be hydroxylated. While HIF-1 is stabilized and localized to the nucleus in both cases, there are differences in transcriptional outcomes between stable hydroxylated and unhydroxylated states66,67.

Additionally, we did not explore the effects of alternative genetic activators of the hypoxic response such as PHD/EGL mutants, which may provide further insight into how different manipulations of the hypoxic response impact longevity.

The circuit-mapping approaches employed in this work are also impacted by limitations in cell-specific genetic modifications and in the use of RNAi knockdown. For example, cell-specific rescue constructs can sometimes lead to unintended rescues in other cell types due to cell-nonautonomous signaling. Because all serotonin-producing neurons also express the serotonin reuptake transporter mod-5, serotonin produced by one cell in our rescue strains could be taken up by other serotonin-producing neurons, leading to unintended signaling effects. This may also be true of the uv1 and RIM tyraminergic rescue strains, although little is known about tyramine reuptake in C. elegans.

Additionally, although RNAi experiments were validated using sequence confirmation and qPCR, not all RNAi hits were corroborated with genetic knockouts. This leaves open the possibility that RNAi-induced knockdown effects differ from complete genetic ablation, or that production of a given RNAi may modify bacterial metabolism in a way that indirectly modifies the hypoxic response in C. elegans. Finally, while the use of RNAi knockdown and genetic knockouts establishes the necessity of many signals within the hypoxia-mediated longevity circuit, the exact directionality of these signals remains unclear. It is possible that increased, decreased, or pulsatile changes in signaling through these bioamines and neuropeptides are required for the hypoxic response to extend lifespan. Future work in this area could use tools to measure or modify neuronal activity, such as calcium imaging or optogenetics, to begin answering these questions.

While many individual neurosignaling components are essential for the hypoxic response to extend lifespan, their epistasis is unclear. Future work in this area should focus on manipulating various components of this network in a manner that we would expect to mimic hypoxia and extend lifespan. This will allow us to narrow down which signals are necessary only to adapt to hypoxia, and which are sufficient to extend lifespan in a normoxic environment. One notable target for further exploration is the SER-7 expressing RIS neuron, which plays a role in sleep68 and stress resistance69. Moreover, optogenetically increasing RIS activity can extend lifespan under normoxic conditions70.

While our understanding of this circuit is still incomplete, our findings point to several promising signaling components that could be targeted for longevity interventions. Manipulating serotonin and tyramine/adrenaline signaling pathways may offer strategies for extending lifespan, with minimal pleiotropic effects if precisely controlled. These neural targets have great potential for longevity therapeutics due to 1) the small number of neurons that can control aging-related pathways in the entire organism; 2) the availability of FDA-approved pharmaceuticals that target individual bioaminergic transporters and receptors71–73; and 3) the high conservation of neurotransmitter biology between invertebrates and mammals50,62,74. Further research into how these pathways can be modulated in a targeted manner could spur development of interventions that promote healthy aging and delay the onset of age-related diseases.

Methods

Strains and Growth Conditions

C. elegans were cultured according to established protocols31. In brief, worms were grown at 20 °C on standard solid nematode growth media (NGM). Throughout their lifespan, the worms were fed E. coli OP50, except during RNA interference (RNAi) experiments, where E. coli HT115 was used to deliver double-stranded RNA. Transfers of worms were carried out using a platinum wire, unless stated otherwise. The RNAi strains employed are listed in Supplementary Table 1, while the strains used in the experiments are provided in Supplementary Table 2. Genotypes were verified through PCR, and RNAi imaging results were confirmed by sequencing and quantitative PCR (qPCR) before proceeding with the experiments.

Generating transgenic strains

Neuron-specific stabilized HIF-1 strains

We used the following promoters to drive cDNA of HIF-1 (P621A)::SL2::GFP in three serotonergic neuronal populations: ADF neurons (srh-142p:: HIF-1 (P621A)), NSM neurons (ceh-2p:: HIF-1 (P621A)), HSN neurons (ham-2p:: HIF-1 (P621A)). All plasmids were verified via restriction digest and sanger sequencing, and ApE files are available upon request.

Plasmids were microinjected to hif-1(ia4) KO strain by Suny Biotech using the co-injection marker myo-2p::GFP and 2-3 transgenic lines were tested in each experiment.

Neuron-specific tph-1 rescues

The pKA805[srh-142p::TPH-1] and pKA807[ceh-2p::TPH-1] constructs were generously provided by Dr. Kaveh Ashrafi. These constructs were injected at ∼50 ng/µL) with fluorescent co-injection marker myo-2p::mNeonGreen (15 ng/µL) or sur-5p::sur-5::NLSGFP (20 ng/µL) and junk DNA (up to 100 ng/µL) into gonads of day 1 gravid adult hermaphrodites. Standard protocols were followed to isolate and obtain stable over-expression mutants75.

ser-7 rescue strains

Plasmid construction and microinjection was conducted by Suny Biotech to generate all ser-7 rescue constructs. In designing the plasmids, we used the following promoters to drive cDNA of SER-7::SL2::GFP in different neuronal populations: full rescue (ser-7p::ser-7), interneuron rescue (glr-1p::ser-7), bioaminergic neuron rescue (cat-1p::ser-7), glutamatergic neuron rescue (eat-4p::ser-7), GABAergic neuron rescue (unc-47p::ser-7), GABAergic motor neuron rescue(unc-25p::ser-7), sensory neuron rescue (osm-6p::ser-7), cholinergic motor neuron rescue(acr-2p::ser-7), cholinergic neuron rescue (unc-17p::ser-7), M3 & M4 neuron rescue (ceh-28p::ser-7), and intestinal rescue (vha-6p::ser-7). All plasmids were verified via restriction digest and sanger sequencing, and ApE files are available upon request. Plasmids were microinjected by Suny Biotech using the co-injection marker myo-2p::GFP and 2-3 transgenic lines were tested in each experiment.

RIS, RIC, and BAG neuronal ablation strains

For RIS neuronal ablation strains, we purchased donor plasmid mec-18p::caspase-3 (p12)::nz [TU#813] (Plasmid #16082) and mec-18p cz::caspase-3 (p17) [TU#814] from Addgene (Plasmid #16083), and used Gibson cloning (NEB) to replace mec-18p with srsx-18p. Three constructs srsx-18p::caspase-3(p12)::nz, srsx-18p::cz::caspase-3(p17), and srsx-18p::GFP were co-injected with fluorescent co-injection marker myo-3p::GFP (20 ng/µL) into the wild-type N2 strain to generate RIS genetic ablation strains. Similarly, for RIC ablation strain, tbh-1p was constructed into TU#813 and TU#814 to replace mec-18p. Three constructs tbh-1p::caspase-3(p12)::nz, tbh-1p::cz::caspase-3(p17), and tbh-1p::GFP were co-injected with fluorescent co-injection marker myo-3p::GFP (20 ng/µL) into the wild-type N2 strain to generate RIC genetic ablation strains. For BAG ablation strain, gcy-31p was constructed into TU#813 and TU#814 to replace mec-18p. Three constructs gcy-31p::caspase-3(p12)::nz, gcy-31p::cz::caspase-3(p17), and gcy-31p::GFP were co-injected with fluorescent co-injection marker myo-3p::GFP (20 ng/µL) into the wild-type N2 strain to generate BAG genetic ablation strains. All plasmids were verified via restriction digest and sanger sequencing. ApE files available upon request. Plasmid construction and microinjection was conducted by Suny Biotech.

RIM and uv1 tdc-1 rescue strains

Plasmid construction and microinjection was conducted by Suny Biotech to generate both tdc-1 rescue constructs. In designing the plasmids, we used the ocr-4 promoter to drive cDNA of TDC-1::SL2::GFP in the uv1 neuroendocrine cells. To express cDNA of TDC-1::SL2::GFP in the RIML neuron, we used the gcy-13 promoter. All plasmids were verified via restriction digest and sanger sequencing, and ApE files are available upon request.

Plasmids were microinjected by Suny Biotech using the co-injection marker myo-2p::GFP.

RNAi Knockdown

For all RNAi knockdowns, worms were exposed to the RNAi treatment for two generations to achieve optimal knockdown. All RNAi clones were sourced from the Vidal RNAi library. Each RNAi clone was sequence-verified.

Quantitative PCR

For RNAi validation experiments, RNA was isolated from day 1 adult worms that had been grown on RNAi for two generations. For measuring fmo-2 induction in the hif-1(ia4); ADF::HIF-1S strain, worms were grown on OP50, synchronized, and collected at day 1 of adulthood. RNA extraction was performed using the Direct-zol RNA Miniprep Kit (Zymo), and cDNA synthesis was carried out with the Biorad iScript cDNA Synthesis Kit. Gene expression was assessed using the SYBR Green quantitative RT-PCR (qRT-PCR) system (BioRad), with mRNA levels normalized to the housekeeping genes cdc-42 and Y45FD10.4. Primers for RNAi validation were designed to target the 3’ UTR of each gene to avoid amplification of the bacterially produced RNAi. Gene expression was quantified using standard curves.

Lifespan Measurements

Lifespan assays were conducted following previously described protocols31. Briefly, 10-15 gravid adults were transferred to NGM plates for a three-hour timed egg lay, after which they were removed. Once the progeny reached day 1 of adulthood, 60-80 worms were placed on NGM plates containing 33 µL of 150 mM fluorodeoxyuridine (FUdR) and 100 µL of 50 mg/mL ampicillin per 100 mL of NGM. FUdR inhibits progeny development, while ampicillin prevents bacterial contamination. NGM + FUdR + Ampicillin plates were seeded with concentrated bacteria (5x for OP50-fed lifespan assays). At least two plates per strain and condition were used for each lifespan replicate. Worms were considered dead when they failed to respond to a gentle touch with a platinum wire under a dissection microscope. Lifespan data were recorded at least three times a week until all worms were dead. To prevent escape, a barrier of 75 µL of 100 mM palmitic acid (Sigma-Aldrich) dissolved in 100% ethanol was applied along the edges of each lifespan plate. Data were analyzed using R version 4.3.1 and visualized in Adobe Illustrator 2022.

RNAi Lifespans

RNAi lifespan assays were performed similarly to standard lifespan assays, with modifications to the initial timed egg lay (TEL) procedure and food concentrations. To ensure maximal RNAi knockdown, worms were first TEL’ed for 3 hours on RNAi plates. The progeny from this TEL were left on RNAi plates to develop into gravid adults and then used for a second TEL under the same RNAi conditions. Progeny from the second-generation TEL were then used for the lifespan assay. RNAi bacteria (HT115) are ampicillin-resistant and can grow slowly on AMP-containing lifespan plates. To maintain consistent bacterial availability throughout the lifespan assay, RNAi plates were seeded with bacteria at 2x concentration, starting from an optical density (OD600) of 3.0.

Fluorescent Slide Microscopy

Fluorescent images for this study were captured using a Leica M165F fluorescent microscope, controlled by Leica Application Suite X (LASX) software. A minimum of 15 worms per condition were imaged at a magnification of at least 70x. For image quantification, individual worms were carefully separated to ensure they did not touch. Custom R code76 was then used to create a pixel mask of each worm from the brightfield image and measure the fluorescent intensity of that region in the corresponding fluorescent image. The background fluorescence was subtracted from the mean fluorescence values. Data were analyzed using R version 4.3.1 and visualized in Adobe Illustrator 2022.

Statistical Analysis

The height of all bar plots represents the mean value for each condition, with error bars indicating the standard error of the mean (SEM). For comparisons involving more than two conditions, one-way ANOVA followed by post-hoc Tukey HSD tests (2-tailed, unpaired) were used to assess interactions between variables and determine statistical differences between conditions. Significance levels are indicated as *p < 0.05, **p < 0.01, ***p < 0.001, and ****p < 0.0001. For lifespan assays, statistical comparisons of survivorship curves were performed using the survfit survival analysis function in R. The log-rank test was used to compare two survival curves. To assess interactions across multiple experimental variables and for comparisons involving more than two survival curves, a Cox regression analysis was conducted using the survivalMPL package in R. For experiments with multiple biological replicates, p-values were adjusted using the Bonferroni correction. Exact sample sizes (Ns) for each experiment are provided in the source data files, and minimum sample sizes for each plot are specified in the figure legends.

eLife Assessment

This fundamental study identifies specific neural mechanisms through which HIF-1 signaling in ADF serotonergic neurons extends lifespan in C. elegans, revealing that downstream signaling in multiple types of neurons, as well as other neuromodulators like GABA, tyramine, and NLP-17, is required for this effect. The strength of the evidence is largely convincing, as the authors establish the necessity and causality of key neuronal components using multiple genetic tools and functional dissection in a well-validated model organism.

Reviewer #1 (Public review):

Summary:

In this study by Kitto et al., the authors set out to identify specific signaling components regulating the hypoxic response from the neurons to the periphery and which components are required for lifespan extension. Their previous work had shown that expression of a stabilized HIF-1 mutant in the nervous system extends lifespan through the serotonin receptor SER-7 and leads to the induction of fmo-2 in the intestine. In the current study, they mapped the precise neural circuits required for this response, as well as the signaling mediators. Their work reveals that neurotransmitters GABA and tyramine, and the neuropeptide NLP-17, act downstream of neuronal HIF-1 to convey a "hypoxic signal" to peripheral tissues. Through cell-type-specific expression studies, targeted knockouts, and comprehensive lifespan analysis, the authors provide robust evidence to support their conclusions. The insights gained from the study are both moving the field forward as they advance our understanding of neuro-peripheral hypoxic signaling, but they also lay the groundwork for potential therapeutic strategies aimed at the modulation of such signaling pathways.

Strengths:

(1) This study provides new evidence further delineating signaling components required for hypoxic signaling-mediated longevity, from the nervous system to the periphery. Using a rigorous approach where they express stabilized HIF-1 mutant selectively in ADF, NSM, and HSN serotonergic neurons, followed by cell-type-specific tph-1 knockouts to pinpoint ADF-dependent serotonin signaling as essential for both lifespan extension and intestinal fmo-2 induction.

This was followed by generating 11 transgenic lines that drive SER-7 expression under distinct neuron-specific promoters, to systematically tease out in which of 27 candidate neurons SER-7 functions to mediate hypoxia-induced longevity. This ultimately highlighted the RIS interneuron as the required signaling hub.

(2) As the intestine lacks direct neuronal innervation, the authors employ neuron-specific RNAi (TU3311 strain) and dense core vesicle analyses to identify that the neuropeptide NLP-17 is required to transmit the hypoxic signal from RIS to induce fmo-2 in the intestine.

(3) Overall, the paper is very well written. The experiments were carried out carefully and thoroughly, and the conclusions drawn are also well supported by the results they are showing.

Weaknesses:

Overall, I don't see many weaknesses. One point relates to their read-outs, which rely heavily on lifespan measurements and fmo-2 induction without evaluating other physiological processes that serotonin or NLP-17 might affect. For translational relevance, it would be valuable to assess or mention potential adverse effects, such as changes in reproduction, pharyngeal pumping, or proteostasis capacity (proteostasis capacity specifically in the tissue showing fmo-2 upregulation).

While lifespan assays and fmo-2 expression do provide strong evidence, incorporating additional markers of stress resistance could strengthen the link between hypoxic signaling and organismal health as well.

Reviewer #2 (Public review):

Summary:

The authors aimed to identify the specific neurons, neurotransmitters, and neuropeptides that mediate the longevity effects of the hypoxic response in C. elegans. By genetically dissecting the pathway downstream of HIF-1, they define a neural circuit involving ADF serotonergic neurons, the SER-7 receptor in the RIS interneuron, tyraminergic signaling from RIM, and neuropeptide NLP-17, ultimately linking neuronal hypoxic sensing to pro-longevity signaling in the intestine.

Strengths:

The study employs a diverse genetic toolkit, including neuron-specific transgenes, tissue-specific knockouts and rescues, RNAi knockdowns, allowing the authors to pinpoint causality, sufficiency, and necessity with high resolution. The comprehensive mapping of cell-nonautonomous signaling adds depth to our understanding of how HIF and serotonin signaling interface with aging pathways. The conclusions are supported by consistent survival assays and fmo-2 gene expression analyses.

Weaknesses:

A key limitation is the lack of clear evidence showing epistasis of so many identified molecular/neuronal components downstream of HIF-1 and serotonin. Thus, the mechanisms of how a diverse set of molecules/neurons coordinate and mediate neuronal HIF-1 effects on intestinal fmo-2 and longevity remain murky. Some rescue strategies may inadvertently cause non-physiological expression. Additionally, environmental hypoxia was not tested in parallel, so the claim on "hypoxia respone" throughout the manuscript is not justified by genetic manipulation alone, and the translational relevance of the genetic manipulations remains somewhat uncertain.

Reviewer #3 (Public review):

Summary:

This study found that ADF serotonergic neurons have a significant role in extending lifespan mediated by HIF-1, as well as serotonin receptor SER-7 in the GABAergic RIS interneurons. The author focuses on the sufficiency and necessity of components from the central nervous system and how they contribute to aging upon hypoxia.

Previous work from the lab has identified that the stabilization of HIF-1 in neurons is sufficient to extend lifespan through the serotonin receptor, SER-7, which subsequently activates fmo-2 in the intestine and leads to lifespan extension. Building on this, the author sought to determine which serotonergic neurons are involved and found that serotonin signaling in ADF neurons is required for lifespan extension mediated by HIF-1.

The author next tested which subset of neurons requires Ser-7 expression to rescue hypoxic response. They found that ser-7 expression in multiple neurons is sufficient to induce fmo-2, with the top candidate being the RIS neuron. Ablation of the RIS neuron did not extend lifespan, suggesting that ser-7 expression in the RIS neuron is required for lifespan extension, positioning it as a key component in the longevity signaling pathway.

The author also investigated neurotransmitters and found that GABA and tyramine are important components in this circuit. They showed that the tyramine receptor called tyra-3 is required for vhl-1-mediated longevity. Given that tyra-3 is expressed in oxygen- and carbon dioxide-sensing neurons, the author demonstrated that these sensing neurons work downstream of serotonin signaling. Lastly, the author screened neuropeptide/receptor binding pairs and identified NLP-17 as playing a role in hypoxia-mediated longevity.

Originality and Significance:

This research is significant in that it uncovers components that are sufficient and necessary for lifespan extension via the hypoxic response. It provides comprehensive data supporting longevity induced by HIF-1-mediated hypoxic response, in conjunction with fmo-2, a longevity gene, as demonstrated in previous work from the lab. Moreover, it provides a number of new transgenic worm tools for C. elegans and aging communities.

Data and Methodology:

(1) The experiments were thoroughly conducted, especially the generations of strains using different neuron-type promoters and crossing into mutant strains to demonstrate sufficiency and necessity.

(2) Some figure legends from the text do not match what the data show. (Figure 6E, F, G).

(3) The lifespan graph legends are confusing and could use some revamping for better clarification.

Conclusions:

This study provides insights into how hypoxic response regulates aging in a cell non-autonomous manner, outlining a potential circuit involving neurons, neurotransmitters, and neuropeptides.

Author response:

The following is the authors’ response to the original reviews.

We thank the reviewers and editors for their time and valuable input into improving our manuscript. The reviewers recognized the value of this work while identifying places where further explanation and/or additional experiments could strengthen the manuscript. We greatly appreciate this feedback and have addressed reviewer comments through additional experiments and necessary textual edits.

In response to the reviewer comments, our principal data-driven changes to revise the manuscript include: 1) testing how stabilizing HIF-1 in the ADF and NSM neurons affects healthspan; 2) measuring for interactions between HIF-1 stabilization in serotonergic neurons and the mt-UPR; and 3) measuring nlp-17 expression downstream of HIF-1 stabilization in serotonergic neurons. We also made textual changes including: 1) clarifying that our study focuses specifically on the vhl-1-mediated genetic hypoxic response; 2) summarizing which parts of the working model are experimentally validated vs. speculative; and 3) expanding our discussion of future directions needed to understand the epistasis of the many signals acting in this circuit. Our responses to each reviewer comment are below.

Public Reviews:

Reviewer #1 (Public review):

Summary:

In this study by Kitto et al., the authors set out to identify specific signaling components regulating the hypoxic response from the neurons to the periphery and which components are required for lifespan extension. Their previous work had shown that expression of a stabilized HIF-1 mutant in the nervous system extends lifespan through the serotonin receptor SER-7 and leads to the induction of fmo-2 in the intestine. In the current study, they mapped the precise neural circuits required for this response, as well as the signaling mediators. Their work reveals that neurotransmitters GABA and tyramine, and the neuropeptide NLP-17, act downstream of neuronal HIF-1 to convey a "hypoxic signal" to peripheral tissues. Through cell-type-specific expression studies, targeted knockouts, and comprehensive lifespan analysis, the authors provide robust evidence to support their conclusions. The insights gained from the study are both moving the field forward as they advance our understanding of neuro-peripheral hypoxic signaling, but they also lay the groundwork for potential therapeutic strategies aimed at the modulation of such signaling pathways.

We appreciate the reviewer’s positive assessment of the topic and general interest in this work.

Strengths:

(1) This study provides new evidence further delineating signaling components required for hypoxic signaling-mediated longevity, from the nervous system to the periphery. Using a rigorous approach where they express stabilized HIF-1 mutant selectively in ADF, NSM, and HSN serotonergic neurons, followed by cell-type-specific tph-1 knockouts to pinpoint ADF-dependent serotonin signaling as essential for both lifespan extension and intestinal fmo-2 induction.

This was followed by generating 11 transgenic lines that drive SER-7 expression under distinct neuron-specific promoters, to systematically tease out in which of 27 candidate neurons SER-7 functions to mediate hypoxia-induced longevity. This ultimately highlighted the RIS interneuron as the required signaling hub.

(2) As the intestine lacks direct neuronal innervation, the authors employ neuron-specific RNAi (TU3311 strain) and dense core vesicle analyses to identify that the neuropeptide NLP-17 is required to transmit the hypoxic signal from RIS to induce fmo-2 in the intestine.

(3) Overall, the paper is very well written. The experiments were carried out carefully and thoroughly, and the conclusions drawn are also well supported by the results they are showing.

Weaknesses:

Overall, I don't see many weaknesses. One point relates to their read-outs, which rely heavily on lifespan measurements and fmo-2 induction without evaluating other physiological processes that serotonin or NLP-17 might affect. For translational relevance, it would be valuable to assess or mention potential adverse effects, such as changes in reproduction, pharyngeal pumping, or proteostasis capacity (proteostasis capacity specifically in the tissue showing fmo-2 upregulation).

We thank the reviewer for the positive review and fully agree and acknowledge that the primary readouts used in this work, fmo-2 induction and lifespan, may not reflect other important elements of health and physiology. To address this weakness, we have performed three measurements of healthspan (pumping, thrashing, and maximum velocity), in the ADF and NSM HIF-1 stabilized strains at young adulthood and at middle age. We focused on examining these HIF-1 stabilized strains rather than our nlp-17 knockout animals because stabilizing HIF-1 in the NSM or ADF neurons is sufficient to extend lifespan. nlp-17 is necessary, but it remains unclear whether nlp-17 signaling is sufficient to extend lifespan. Our new data show that both ADF- and NSM-specific HIF-1 stabilization had no effect on pumping rate in young (day 1 of adulthood) worms. In aged animals (day 12 of adulthood), however, NSM-, but not ADF-, specific HIF-1 stabilization rescued the pumping rate decline in hif-1 knockout compared to WT worms (new Fig. S1C). Similarly in the new thrashing data, NSM-, but not ADF-, specific HIF-1 stabilization rescued the thrashing rate decline in the hif-1 knockout young and aged worms (new Fig. S1D). Lastly, our new data show that ADF and NSM HIF-1 stabilization had no effect on average or maximum movement speed at days 1 and 5 of adulthood (new Fig. S1E-F). Together, these results indicate that genetic activation of the hypoxic response in NSM neurons but not in the ADF neurons could improve healthspan.

While we did not examine reproductive capacity in these strains, we have expanded our discussion section to mention the importance of fully characterizing other elements of physiology, health, and behavior in future work.

“Another limitation of this work is that it uses lifespan as the main readout for organismal health. While genetic manipulations that extend lifespan often improve stress resistance and healthspan [8,80], longevity manipulations can also have adverse effects on reproduction [81,82] and behavior [83,84]. In this study, we find that HIF-1 stabilization in the ADF neurons does not prevent the deleterious effects of hif-1 knockout on mobility, but that HIF-1 stabilization in the NSM neurons may attenuate age-related decline in pumping and thrashing (Fig. S1). However, future work should examine whether other modifications to this pathway, such as manipulations to RIM, RIS, or NLP-17 signaling, influence healthspan in addition to lifespan. It will be important for future studies to determine whether various components of this pathway affect both longevity and the response to different types of stressors like oxidative stress, proteotoxic stress, and infection, as HIF-1 activity also interacts with multiple stress responses [41,42,39,40,43].”

While lifespan assays and fmo-2 expression do provide strong evidence, incorporating additional markers of stress resistance could strengthen the link between hypoxic signaling and organismal health as well.

We also measured hsp-6 expression via qPCR in the serotonergic neuron-specific HIF-1 stabilized strains to determine whether these conditions that lead to upregulated fmo-2 may also affect the mt-UPR. Interestingly, we find that stabilizing HIF-1 in either the ADF or NSM serotonergic neurons decreases hsp-6 expression relative to WT worms (new Fig. S1G). This could suggest either that the mt-UPR response is impaired in these worms, or that HIF-1 stabilization decreases proteotoxic stress leading to a lower basal level of hsp-6. Although this method of measurement did not allow us to interrogate whether these changes occur in the specific tissues where fmo-2 is upregulated, we have expanded our discussion to emphasize that further investigation of cell- and tissue-specificity within the hypoxic response should be a focus of future work.

Finally, we agree that it is important to examine whether activating the hypoxic response promotes stress resistance in addition to longevity. We did not focus on this element of the hypoxic response in this work because HIF activity is known to promote adaptive stress-responses to some stressors like infection [7,8] and oxidative stress [9,10], while simultaneously impairing the proteosasis stress response [11]. We have added this important information to our introduction section, and have expanded our discussion section to emphasize that a key future direction will be to test which specific components of this pathway also facilitate stress resistance.

Introduction Section Modification:

“However, the physiological changes induced by the hypoxic response are broad and involve adaptations such as increased vascularization, metabolic rewiring, and changes in cell survival pathways. In mammals, some of these same adaptations can be detrimental, as mutations in components of the hypoxic response have been linked to conditions like cancer and cardiovascular disease [12-15]. Additionally, HIF activity is essential to promote some forms of stress resistance to infection [9,10] and oxidative stressors [7,8] but can have detrimental effects on proteostasis [11].”

Discussion Section Modification:

“Another limitation of this work is that it uses lifespan as the main readout for organismal health. While genetic manipulations that extend lifespan often improve stress resistance and healthspan [1,2], longevity manipulations can also have adverse effects on reproduction [3,4] and behavior [5,6]. In this study, we find that HIF-1 stabilization in the ADF or NSM neurons has little effect on mobility in young and aged animals. However, future work should examine whether other modifications to this pathway, such as manipulations to RIM, RIS, or NLP-17 signaling, influence healthspan in addition to lifespan. It will be important for future studies to determine whether various components of this pathway affect both longevity and the response to different types of stressors like oxidative stress, proteotoxic stress, and infection, as HIF-1 activity also interacts with multiple stress responses [7,8,9-11].”

Reviewer #2 (Public review):

Summary:

The authors aimed to identify the specific neurons, neurotransmitters, and neuropeptides that mediate the longevity effects of the hypoxic response in C. elegans. By genetically dissecting the pathway downstream of HIF-1, they define a neural circuit involving ADF serotonergic neurons, the SER-7 receptor in the RIS interneuron, tyraminergic signaling from RIM, and neuropeptide NLP-17, ultimately linking neuronal hypoxic sensing to pro-longevity signaling in the intestine.

Strengths:

The study employs a diverse genetic toolkit, including neuron-specific transgenes, tissue-specific knockouts and rescues, RNAi knockdowns, allowing the authors to pinpoint causality, sufficiency, and necessity with high resolution. The comprehensive mapping of cell-nonautonomous signaling adds depth to our understanding of how HIF and serotonin signaling interface with aging pathways. The conclusions are supported by consistent survival assays and fmo-2 gene expression analyses.

Weaknesses:

A key limitation is the lack of clear evidence showing epistasis of so many identified molecular/neuronal components downstream of HIF-1 and serotonin. Thus, the mechanisms of how a diverse set of molecules/neurons coordinate and mediate neuronal HIF-1 effects on intestinal fmo-2 and longevity remain murky.

We thank the reviewer for these important points. We agree that the epistatic relationships between ADF serotonin, RIM tyramine, RIS GABA, and neuropeptide NLP-17 signaling remain unclear within this pathway. Determining the epistatic relationships of each signal within this complex pathway will require: 1) generating genetic manipulations to each identified signaling component that may mimic vhl-1 knockout to promote longevity; 2) crossing these new strains into multiple genetic knockouts we identified as required for vhl-1 mediated longevity; and 3) measuring the lifespans of each double and triple mutant. We are very interested in testing the epistatic relationships between all of these molecules and neurons, and believe this extensive follow-up exploration will generate significant future results.

To better address these limitations of the current work, we have added a summary table to Fig. 7 as well as a paragraph to our discussion section. This table and paragraph better explain which components of our working model have been tested for necessity, sufficiency, and epistasis, and which components of this model remain unclear (Fig. 7). See updated Fig. 7 with added table.

Updated discussion section detailing this limitation:

“While many individual neurosignaling components are essential for genetic activation of the hypoxic response to extend lifespan, their epistasis is unclear (Fig. 7B). Most components of the pathway identified in this work act downstream of vhl-1 depletion, and upstream of fmo-2 induction (summarized in Fig. 7A-B). However, the order of each signal between these two endpoints is only predicted based on C. elegans neural wiring and the overlap between various identified signals and cells. For example, we hypothesize in our working model that GABA may be produced by the RIS neuron in this circuit because RIS is the primary GABAergic neuron required for vhl-1-mediated longevity. Alternatively, it is possible that GABA is produced by a different cell that either acts in series or in parallel with RIS signaling. In order to determine the order of each signaling component, future studies should generate genetic manipulations to each signaling component that may mimic vhl-1 knockout to promote longevity, cross these new strains into knockouts of other signals required for vhl-1 mediated longevity; and measure the lifespans of each double and triple mutant. This approach would also narrow down which signals are downstream of the genetic activation of the hypoxic response, and which are sufficient to extend lifespan upstream of the hypoxic response in a normoxic environment. One notable target for further exploration is the SER-7 expressing RIS neuron, which plays a role in sleep [16] and stress resistance [17], and can extend lifespan when optogenetically activated under normoxic conditions [18].”

Some rescue strategies may inadvertently cause non-physiological expression.

This is a great point. We bring attention to this limitation in the discussion section. Our cell-specific knockouts (ADF tph-1 KO, Fig. 1E) or ablations (RIS ablation, Fig. 2C) data showed that the neurons identified using rescue strains (i.e., tph-1 in the ADF and ser-7 in RIS) are likely not false positives. However, we did not generate a RIM-specific knockout or ablation strain to confirm our tdc-1 results and have added this limitation to the discussion of these results.

Discussion of rescue strategy limitations and the need for a RIM-specific knockout:

“The circuit-mapping approaches employed in this work are also impacted by limitations in cell-specific genetic modifications and in the use of RNAi knockdown. For example, cell-specific rescue constructs can sometimes lead to unintended rescues in other cell types due to cell-nonautonomous signaling. Because all serotonin-producing neurons also express the serotonin reuptake transporter mod-5, serotonin produced by one cell in our rescue strains could be taken up by other serotonin-producing neurons, leading to unintended signaling effects. This may also be true of the uv1 and RIM tyraminergic rescue strains, although little is known about tyramine reuptake in C. elegans. Two cells identified via tissue-specific rescue experiments (ADF and RIS, Fig. 1G and Fig. 2D) were also found to be necessary via cell-specific knockout (ADF tph-1 KO, Fig. 1E; RIS ablation, Fig. 2C), decreasing the likelihood of a false positive from the rescue strain technique. However, the role of the RIM neuron was identified via a tdc-1 rescue strain and was not validated using a RIM-specific knockout (Fig. 7B). Therefore, the contribution of RIM signaling to this circuit is less well-validated, and a RIM-specific tdc-1 knockout strain should be examined in future work.”

Additionally, environmental hypoxia was not tested in parallel, so the claim on "hypoxia response" throughout the manuscript is not justified by genetic manipulation alone, and the translational relevance of the genetic manipulations remains somewhat uncertain.

We thank the reviewer for identifying the need for additional specificity in our language. We agree that it is critical to make it clear that this paper only examines genetic mimics of hypoxia, rather than environmental hypoxia, and have replaced every occurrence of “the hypoxic response” with either 1) “genetic activation of the hypoxic response”, or 2) describing the specific manipulation used in that experiment (e.g., “vhl-1 mediated longevity”). We also agree that examining the similarities and differences between genetic and environmental activation is a key next step towards evaluating the translational potential of this pathway. We discuss this in Discussion and are excited to interrogate these differences in future work.

“While this study identifies many neural signals required for vhl-1 knockdown or knockout to extend lifespan, one key limitation of this work is the potential differences between genetic and environmental methods of inducing the hypoxic response. While vhl-1 knockdown or knockout leads to HIF-1 stabilization by blocking its proteasomal degradation, it also results in hydroxylated HIF-1. This contrasts with environmental hypoxia, in which HIF-1 remains stable because it cannot be hydroxylated. While HIF-1 is stabilized and localized to the nucleus in both cases, there are differences in transcriptional outcomes between stable hydroxylated and unhydroxylated states [19,20]. Additionally, we did not explore the effects of alternative genetic activators of the hypoxic response such as HIF-1 hydroxylase PHD/EGL mutants, which may provide further insight into how different manipulations of the hypoxic response impact longevity. Future work should interrogate the similarities and differences between the circuits driving longevity in response to environmental hypoxia, vhl-1 knockdown, HIF-1 stabilization, and PHD/EGL knockdown.”

Reviewer #3 (Public review):

Summary:

This study found that ADF serotonergic neurons have a significant role in extending lifespan mediated by HIF-1, as well as serotonin receptor SER-7 in the GABAergic RIS interneurons. The author focuses on the sufficiency and necessity of components from the central nervous system and how they contribute to aging upon hypoxia.

Previous work from the lab has identified that the stabilization of HIF-1 in neurons is sufficient to extend lifespan through the serotonin receptor, SER-7, which subsequently activates fmo-2 in the intestine and leads to lifespan extension. Building on this, the author sought to determine which serotonergic neurons are involved and found that serotonin signaling in ADF neurons is required for lifespan extension mediated by HIF-1.

The author next tested which subset of neurons requires Ser-7 expression to rescue hypoxic response. They found that ser-7 expression in multiple neurons is sufficient to induce fmo-2, with the top candidate being the RIS neuron. Ablation of the RIS neuron did not extend lifespan, suggesting that ser-7 expression in the RIS neuron is required for lifespan extension, positioning it as a key component in the longevity signaling pathway.

The author also investigated neurotransmitters and found that GABA and tyramine are important components in this circuit. They showed that the tyramine receptor called tyra-3 is required for vhl-1-mediated longevity. Given that tyra-3 is expressed in oxygen- and carbon dioxide-sensing neurons, the author demonstrated that these sensing neurons work downstream of serotonin signaling. Lastly, the author screened neuropeptide/receptor binding pairs and identified NLP-17 as playing a role in hypoxia-mediated longevity.

Originality and Significance:

This research is significant in that it uncovers components that are sufficient and necessary for lifespan extension via the hypoxic response. It provides comprehensive data supporting longevity induced by HIF-1-mediated hypoxic response, in conjunction with fmo-2, a longevity gene, as demonstrated in previous work from the lab. Moreover, it provides a number of new transgenic worm tools for C. elegans and aging communities.

We thank the reviewer for the positive assessment of the manuscript. We appreciate all suggestions for further improving the work and have made changes based on these suggestions (see details below).

Data and Methodology:

(1) The experiments were thoroughly conducted, especially the generations of strains using different neuron-type promoters and crossing into mutant strains to demonstrate sufficiency and necessity.

(2) Some figure legends from the text do not match what the data show. (Figure 6E, F, G).

We have made changes to the legends accordingly to make sure figures and legends are consistent.

(3) The lifespan graph legends are confusing and could use some revamping for better clarification.

We have updated the lifespan graph legends to provide the statistics in the same section as the labels, indicating which line is which condition (see updated Figure 1E). We hope this change better clarifies the lifespan graph legends.

Conclusions:

This study provides insights into how hypoxic response regulates aging in a cell non-autonomous manner, outlining a potential circuit involving neurons, neurotransmitters, and neuropeptides.

Recommendations for the authors:

Reviewing Editor Comments:

As suggested by the Reviewers 2 & 3, including environmental hypoxia will broaden the impact of the study. If not, the authors should consider clarifying their response as "genetic activation of the hypoxic response" (see Reviewer 2 below).

We thank the editor and reviewers 2 and 3 for this clarification. We agree that it is critical to make it clear that this paper only examines genetic mimics of hypoxia, rather than environmental hypoxia, and have replaced every occurrence of “the hypoxic response” with either 1) “genetic activation of the hypoxic response”, or 2) describing the specific manipulation used in that experiment (e.g., “vhl-1 mediated longevity”). We also agree that examining the similarities and differences between genetic and environmental activation is a key next step towards evaluating the translational potential of this pathway. We have also expanded our discussion of this important future direction.

“While this study identifies many neural signals required for vhl-1 knockdown or knockout to extend lifespan, one key limitation of this work is the potential differences between genetic and environmental methods of inducing the hypoxic response. While vhl-1 knockdown or knockout leads to HIF-1 stabilization by blocking its proteasomal degradation, it also results in hydroxylated HIF-1. This contrasts with environmental hypoxia, in which HIF-1 remains stable because it cannot be hydroxylated. While HIF-1 is stabilized and localized to the nucleus in both cases, there are differences in transcriptional outcomes between stable hydroxylated and unhydroxylated states [19,20]. Additionally, we did not explore the effects of alternative genetic activators of the hypoxic response such as HIF-1 hydroxylase PHD/EGL mutants, which may provide further insight into how different manipulations of the hypoxic response impact longevity. Future work should interrogate the similarities and differences between the circuits driving longevity in response to environmental hypoxia, vhl-1 knockdown, HIF-1 stabilization, and PHD/EGL knockdown.”

Suggested minor changes by Reviewer 3 should also be made to improve the readability of the paper.

We have made changes suggested by Reviewer 3 to improve the readability of the paper.

Reviewer #2 (Recommendations for the authors):

(1) Suggestions for additional experiments or analyses:

(a) To clarify the hierarchical relationships among the components of the identified circuit, epistasis experiments between serotonin, tyramine, NLP-17, and oxygen-sensing neurons (e.g., double mutants or sequential rescues) would strengthen the proposed model and help determine whether these signals act in parallel or downstream of each other.

We thank the reviewer for identifying this important caveat. As described in the public review response, we completely agree that understanding the epistasis of serotonin, tyramine, NLP-17, and oxygen-sensing neuron signaling within this pathway is important to fully test our working model. We hope to address these questions about epistasis and interactions between different signals in upcoming projects.

To clarify that the order of many of these signals remains speculative in our working model, we have added a table to Fig. 7, that summarizes what is known and unknown about the epistasis of these pathway components. We have also expanded our discussion of this limitation. See updated Fig. 7 with added table.

Updated discussion section detailing this limitation:

“While many individual neurosignaling components are essential for genetic activation of the hypoxic response to extend lifespan, their epistasis is unclear (Fig. 7B). Most components of the pathway identified in this work act downstream of vhl-1 depletion, and upstream of fmo-2 induction (summarized in Fig. 7A-B). However, the order of each signal between these two endpoints is only predicted based on C. elegans neural wiring and the overlap between various identified signals and cells. For example, we hypothesize in our working model that GABA may be produced by the RIS neuron in this circuit because RIS is the primary GABAergic neuron we found to be required for vhl-1-mediated longevity. Alternatively, it is possible that GABA is produced by a different cell that either acts in series or in parallel with RIS signaling. In order to determine the order of each signaling component, future studies should generate genetic manipulations to each signaling component that may mimic vhl-1 knockout to promote longevity, cross these new strains into knockouts of other signals required for vhl-1 mediated longevity; and measure the lifespans of each double and triple mutant.”

(b) Testing whether environmental hypoxia (e.g., 0.5-1% O₂ exposure) elicits similar neuronal requirements and fmo-2 induction as the genetic HIF-1 stabilization would validate that the described pathway is relevant to the actual hypoxic response and improve translational relevance.

We thank the reviewer for identifying the need for clarification. As also discussed in the public review section, we agree that it is important to emphasize that this paper exclusively examines genetic mimetics of hypoxia, rather than actual exposure to a hypoxic environment. To clarify this point, we have replaced every occurrence of “the hypoxic response” with either 1) “genetic activation of the hypoxic response”, or 2) describing the specific manipulation used in that experiment (ie “vhl-1 mediated longevity”). We also agree that examining the similarities and differences between genetic and environmental activation is a key next step towards evaluating the translational potential of this pathway. We discuss this in discussion and are excited to interrogate these differences in future work.

“While this study identifies many neural signals required for vhl-1 knockdown or knockout to extend lifespan, one key limitation of this work is the potential differences between genetic and environmental methods of inducing the hypoxic response. While vhl-1 knockdown or knockout leads to HIF-1 stabilization by blocking its proteasomal degradation, it also results in hydroxylated HIF-1. This contrasts with environmental hypoxia, in which HIF-1 remains stable because it cannot be hydroxylated. While HIF-1 is stabilized and localized to the nucleus in both cases, there are differences in transcriptional outcomes between stable hydroxylated and unhydroxylated states [19,20]. Additionally, we did not explore the effects of alternative genetic activators of the hypoxic response such as HIF-1 hydroxylase PHD/EGL mutants, which may provide further insight into how different manipulations of the hypoxic response impact longevity. Future work should interrogate the similarities and differences between the circuits driving longevity in response to environmental hypoxia, vhl-1 knockdown, HIF-1 stabilization, and PHD/EGL knockdown.”

(c) Functional readouts beyond lifespan and fmo-2 induction (e.g., neuronal activity monitoring or optogenetic modulation of key neurons) could help clarify how information flows through the circuit.

We thank the reviewer for this valuable suggestion. We are hoping to be able to implement these experimental techniques in future work. We agree that in combination with genetic epistasis analyses, direct measurements of neuronal signaling will greatly improve our understanding of the directionality and interactions between signals in this circuit. Our discussion section recommends employing these techniques in future studies.

“Finally, while the use of RNAi knockdown and genetic knockouts establishes the necessity of many signals within the vhl-1-mediated longevity circuit, the exact directionality of these signals remains unclear. It is possible that increased, decreased, or pulsatile changes in signaling through these bioamines and neuropeptides are required for genetic activation of the hypoxic response to extend lifespan. Work on C. elegans reversal behavior has also revealed an antagonistic relationship between RIM and RIS activity facilitated by both chemical (neuropeptide and tyramine) and electrical (gap junction) signaling [16,21]. This known interaction should also be interrogated in the context of how these cells may communicate following genetic induction of the hypoxic response. Future work in this area could use tools to measure or modify neuronal activity, such as calcium imaging or optogenetics, to begin answering these questions.”

(2) Recommendations for improving the writing and presentation:

(a) The manuscript is well written overall, but clarity would be improved by explicitly stating in the abstract and introduction that the study is based on genetic activation of the hypoxic response, rather than environmental hypoxia.

We appreciate this valuable suggestion and have updated the abstract and introduction to clarify that this work examines genetic activation of the hypoxic response.

Updated sentences from the abstract:

“Here, we interrogate the cell-nonautonomous signaling pathway downstream of genetic activation of the hypoxic response.”

“Together, these insights develop a circuit for how genetic induction of the hypoxic response cell-nonautonomously modulates ageing and suggests valuable targets for modulating ageing in mammals.”

Updated sentences from the introduction:

“In this study, we uncover key neural components of the longevity circuit initiated by genetic induction of the hypoxic response. Within this circuit, we identify individual cells, signals, and receptors necessary and/or sufficient to extend lifespan downstream of genetic activation of the hypoxic response. More specifically, we find serotonin signaling in the ADF serotonergic neurons is both necessary and sufficient to extend lifespan through genetic activation of the hypoxic response. This pathway signals through the serotonin receptor SER-7 in the RIS interneuron. We further demonstrate additional neurotransmitters (GABA and tyramine), and a neuropeptide (NLP-17) are critical for mediating these longevity effects. Finally, we identify that oxygen sensing neurons (URX, AQR, PQR and BAG) act downstream of neuronal HIF-1 in this circuit. Our insights into this longevity pathway provide a mechanistic understanding of how genetic activation of the hypoxic response delays aging and improves health.”

(b) In the discussion, clearly delineating which parts of the proposed pathway are firmly established versus inferred would aid interpretation.

We thank the reviewer for this idea, and have added the following text to the discussion:

“Evidence for the necessity, sufficiency, and epistatic relationships between each signal are summarized in new Fig. 7B. In brief, all signaling molecules presented in this work are necessary for vhl-1 to extend lifespan. Rescuing ADF serotonin production, RIS ser-7 expression, and RIM tyramine production is sufficient for vhl-1 to extend lifespan. The sufficiency of oxygen sensing neurons BAG and UPA/PQR/AQR, and the neuronal signals of GABA, NLP-17, and TYRA-3 to restore vhl-1 mediated longevity remains unclear. All signals act downstream of vhl-1. ADF HIF-1 stabilization and SER-7 signaling act upstream of fmo-2 induction, and the oxygen sensing neurons (BAG, UPA/PQR/AQR) act upstream of or in parallel to ADF HIF-1 stabilization.”

(c) Adding a summary table or schematic that visually distinguishes necessity vs. sufficiency for each component (e.g., ADF, RIS, RIM, NLP-17) would make the overall model more accessible.

We thank the reviewer for this great suggestion and have added a table to new Fig. 7B that summarizes what is known about necessity for vhl-1, sufficiency for vhl-1, and sufficiency to extend lifespan independently of vhl-1 for each signal (table included in response to Reviewer 2, comment 1a).

(3) Minor corrections and clarifications:

(a) Define or replace "hypoxic response" with "genetically induced hypoxic response" where appropriate to avoid conflating genetic manipulations with actual environmental hypoxia.

We appreciate this valuable suggestion and have replaced “the hypoxic response” and with “genetic activation of the hypoxic response” or “genetically induced hypoxic response” throughout the manuscript to clarify this point.

(b) All genes and alleles should be italicized per worm nomenclature.

We thank the reviewer for this comment and have reviewed the manuscript to italicize all gene names and alleles. In some locations, the protein is referred to instead of the gene using the conventional uppercase non-italicized format.

Reviewer #3 (Recommendations for the authors):

(1) Using vhl-1 RNAi as the sole approach to demonstrate hypoxic response appears somewhat limited, as vhl-1 is also involved in HIF-1 independent processes that can influence lifespan in C. elegans. Including additional downstream effectors of HIF-1, such as egl-9, or having HIF-1 nondegradable strain as validation could strengthen the findings.

We thank the reviewer for this valuable comment and agree that an important next step is to test whether these signals are also required for other genetic (HIF-1 stabilized, egl-9) and environmental activators of the hypoxic response to extend lifespan. We have worked to clarify that this paper focuses primarily on vhl-1 mediated longevity throughout the text and have also included this limitation in our discussion section (for details, please see response to Reviewer 2 public review).

(2) Investigating how the healthspan is affected by serotonergic neuron-specific hypoxic responses would be interesting and could enhance understanding of the physiological mechanisms underlying lifespan extension.

We appreciate this suggestion, and have performed three measurements of healthspan (pumping, thrashing, and maximum velocity), in the ADF and NSM HIF-1 stabilized strains at young adulthood and at middle age. We found that both ADF- and NSM-specific HIF-1 stabilization had no effect on pumping rate in young (day 1 of adulthood) worms. In aged animals (day 12 of adulthood), however, NSM-, but not ADF-, specific HIF-1 stabilization rescued the pumping rate decline in hif-1 knockout compared to WT worms (new Fig. S1C). Similarly, NSM-, but not ADF-, specific HIF-1 stabilization rescued the thrashing rate decline in the hif-1 knockout young and aged worms (new Fig. S1D). ADF and NSM HIF-1 stabilization also had no effect on average or maximum movement speed at days 1 and 5 of adulthood (new Fig. S1E-F). Together, these results indicate that genetic activation of the hypoxic response in NSM neurons but not in the ADF neurons could improve healthspan.

(3) While the experiments were thoroughly performed, the connections between components such as NLP-17, GABA, and tyramine in regulating aging appear critical for establishing a "cell non-autonomous circuit." Additionally, how the potentially antagonistic roles of RIM and RIS neurons influence this axis could be interesting to further explore.

We thank the reviewer for identifying this important caveat. As described in the public review response to Reviewer 2, we completely agree that understanding the epistasis of serotonin, tyramine, NLP-17, and oxygen-sensing neuron signaling within this pathway is important to fully test our working model. Our current data showed that intestinal fmo-2 is required for neuronal HIF-1 stabilization to extend lifespan, indicating information must be communicated between the nervous system and the intestine through serotonin, tyramine, NLP-17 and responsible neurons using a “cell non-autonomous circuit” [22]. However, we will continue to address questions about epistasis and interactions between different signals in upcoming projects to fully establish the circuit.

With respect to RIM and RIS, we value this suggestion and agree that there could be interesting signaling occurring between RIM and RIS in this circuit, as is observed in initiation of reversal behaviors [16,21]. We have updated the discussion section to mention this interesting antagonistic relationship between RIM and RIS signaling in the context of reversal behaviors:

“Finally, while the use of RNAi knockdown and genetic knockouts establishes the necessity of many signals within the vhl-1-mediated longevity circuit, the exact directionality of these signals remains unclear. It is possible that increased, decreased, or pulsatile changes in signaling through these bioamines and neuropeptides are required for genetic activation of the hypoxic response to extend lifespan. Work on C. elegans reversal behavior has also revealed an antagonistic relationship between RIM and RIS activity facilitated by both chemical (neuropeptide and tyramine) and electrical (gap junction) signaling [16,21]. This known interaction should also be interrogated in the context of how these cells may communicate following genetic induction of the hypoxic response. Future work in this area could use tools to measure or modify neuronal activity, such as calcium imaging or optogenetics, to begin answering these questions.”

(4) Does serotonergic neuron-specific rescue impact the mitochondrial unfolded protein response (mtUPR), given that serotonin signaling has been shown to modulate mtUPR?

We appreciate this question and suggestion. To determine whether activating the hypoxic response in serotonergic neurons modifies the mt-UPR, we measured hsp-6 expression via qPCR in the ADF and NSM-specific HIF-1 stabilized strains. Interestingly, we find that stabilizing HIF-1 in either the ADF or NSM serotonergic neurons decreases hsp-6 expression relative to WT worms (new Fig. S1G. This could suggest either that the mt-UPR response is impaired in these worms, or that HIF-1 stabilization decreases proteotoxic stress leading to a lower basal level of hsp-6. Although this method of measurement did not allow us to interrogate whether these changes occur in the specific tissues where fmo-2 is upregulated, we have expanded our discussion of these results to emphasize that further investigation of this response should be a focus of future work.

(5) To remain consistent with the flow of Figure 1, the authors should include fmo-2 expression in NSM:HIF-1S in addition to the ADF:HIF-1S (Figure 1I).

We appreciate this suggestion and have added NSM::HIF-1S data to Fig. 1I. We found that stabilizing HIF-1 in either the ADF or the NSM has a similar effect on fmo-2 induction.

(6) It would greatly strengthen the RIS observation if the authors demonstrated that ablation of another neural subtype from their screen does not abolish lifespan extension by vhl-1 RNAi. This would be a good supplemental figure, but it is not necessary for the overall story.

We thank the reviewer for this suggestion and agree that ablating a ser-7 expressing neuron that was not a hit from our screen would be an excellent additional control. We did find that ablating a non-ser-7-expressing neuron (the RIC, Fig. 3E) did not affect vhl-1 mediated longevity, suggesting that impairing the signaling of any interneuron is not sufficient to disrupt the phenotype. In addition, ablating a ser-7 expressing neuron other than RIS would provide much stronger support for this finding. We have suggested this approach to further validate this working model in the future directions section of our discussion.

“Finally, additional genetic controls could better support the role of RIS-specific ser-7 expression in genetic activation of the hypoxic response. For example, a ser-7 expressing neuron that was not a hit in our screen could also be ablated and tested for necessity in vhl-1 mediated longevity. This experiment would test whether the ability of RIS ablation to attenuate vhl-1-mediated longevity is not a false positive driven by any disruption to ser-7 expression.”

(7) For consistency with the rest of the manuscript, it would strengthen the hypothesis if modulating the expression of nlp-17 or its receptors impacted the intestinal activation of fmo-2 transcription.

This is a great point. We attempted this experiment, but were unable to achieve consistent results (see Author response image 1). This result could be due to indirect effects of the overexpression of nlp-17 signaling modulating fmo-2 induction in a complicated circuit, variability in expression of its receptor, or other complexities within the circuit.

(8) It would strengthen the manuscript to determine whether serotonergic, GABA, and/or tyramine signaling activate the expression or secretion of this nlp-17 neuropeptide.

We thank the reviewer for this great idea of experiment. To address this suggestion, we performed qPCR to measure nlp-17 mRNA in WT, hif-1 KO, ADF HIF-1 stabilized, and NSM HIF-1 stabilized strains. Compared to WT and hif-1 KO controls, we observed no change in nlp-17 expression when HIF-1 was stabilized in the ADF or NSM serotonergic neurons (new Fig. S5F). This could suggest either that nlp-17 signaling acts in parallel to serotonergic signaling following genetic activation of the hypoxic response. Alternatively, neuronal HIF-1 stabilization may modify nlp-17 splicing or translation without resulting in detectable differences in mRNA levels. Together, these data indicate NLP-17 signaling is required for longevity following genetic activation of the hypoxic response, although whether this peptide is synthesized or released in response to hypoxic response remains unclear.

We agree that it is also important to connect nlp-17 expression and/or secretion to other components of this pathway. However, we believe the most effective experiment to confirm a connection between GABA and tyramine signaling and nlp-17 in the context of hypoxia would be to measure nlp-17 expression in strains that manipulate GABA and/or tyramine signaling in a manner that mimics vhl-1 knockdown and extends lifespan. Because we have not yet validated hypoxic-response mimetics for these specific signals, we hope to first generate these strains and then measure their effect on nlp-17 expression in future work. The importance of identifying manipulations to GABA and tyramine signaling that promote longevity has been added to our discussion section.

“While many individual neurosignaling components are essential for genetic activation of the hypoxic response to extend lifespan, their epistasis is unclear (Fig. 7B). Most components of the pathway identified in this work act downstream of vhl-1, and upstream of fmo-2 induction (summarized in Fig. 7A-B). However, the order of each signal between these two endpoints is only predicted based on C. elegans neural wiring and the overlap between various identified signals and cells. For example, we hypothesize in our working model that GABA may be produced by the RIS neuron in this circuit because RIS is the primary GABAergic neuron required for vhl-1-mediated longevity. Alternatively, it is possible that GABA is produced by a different cell that either acts in series or in parallel with RIS signaling. In order to determine the order of each signaling component, future studies should generate genetic manipulations to each signaling component that may mimic vhl-1 knockout to promote longevity, cross these new strains into knockouts of other signals required for vhl-1 mediated longevity; and measure the lifespans of each double and triple mutant. This approach would also narrow down which signals are downstream of the genetic activation of the hypoxic response, and which are sufficient to extend lifespan upstream of the hypoxic response in a normoxic environment. One notable target for further exploration is the SER-7 expressing RIS neuron, which plays a role in sleep [16] and stress resistance [17], and can extend lifespan when optogenetically activated under normoxic conditions [18].”

References:

(1) Calabrese, E. J., Dhawan, G., Kapoor, R., Iavicoli, I. & Calabrese, V. What is hormesis and its relevance to healthy aging and longevity? Biogerontology 16, 693-707 (2015). https://doi.org/10.1007/s10522-015-9601-0

(2) Zhou, I. K., Pincus, Z. & Slack, J. F. Longevity and stress in Caenorhabditis elegans. Aging 3, 733-753 (2011). https://doi.org/10.18632/aging.100367

(3) Yuan, R., Hascup, E., Hascup, K. & Bartke, A. Relationships among Development, Growth, Body Size, Reproduction, Aging, and Longevity - Trade-Offs and Pace-Of-Life. Biochemistry (Mosc) 88, 1692-1703 (2023). https://doi.org/10.1134/S0006297923110020

(4) Mautz, B. S., Lind, M. I. & Maklakov, A. A. Dietary Restriction Improves Fitness of Aging Parents But Reduces Fitness of Their Offspring in Nematodes. J Gerontol A Biol Sci Med Sci 75, 843-848 (2020). https://doi.org/10.1093/gerona/glz276

(5) Duric, V., Clayton, S., Leong, L. M. & Yuan, L.-L. Comorbidity Factors and Brain Mechanisms Linking Chronic Stress and Systemic Illness. Neural Plasticity 2016, 1-16 (2016). https://doi.org/10.1155/2016/5460732

(6) Mariotti, A. The Effects of Chronic Stress On Health: New Insights Into the Molecular Mechanisms of Brain–Body Communication. Future Science OA 1 (2015). https://doi.org/10.4155/fso.15.21

(7) Bellier, A., Chen, C.-S., Kao, C.-Y., Cinar, H. N. & Aroian, R. V. Hypoxia and the Hypoxic Response Pathway Protect against Pore-Forming Toxins in C. elegans. PLoS Pathog 5, e1000689 (2009).

(8) Palazon, A., Goldrath, W. A., Nizet, V. & Johnson, S. R. HIF Transcription Factors, Inflammation, and Immunity. Immunity 41, 518-528 (2014). https://doi.org/10.1016/j.immuni.2014.09.008

(9) Vora, M. et al. The hypoxia response pathway promotes PEP carboxykinase and gluconeogenesis in C. elegans. Nature Communications 13 (2022). https://doi.org/10.1038/s41467-022-33849-x

(10) Nakazawa, S. M., Keith, B. & Simon, C. M. Oxygen availability and metabolic adaptations. Nature Reviews Cancer 16, 663-673 (2016). https://doi.org/10.1038/nrc.2016.84

(11) Fawcett, M. E., Hoyt, M. J., Johnson, K. J. & Miller, L. D. Hypoxia disrupts proteostasis in Caenorhabditis elegans. Aging Cell 14, 92-101 (2015). https://doi.org/10.1111/acel.12301

(12) Ohh, M., Taber, C. C., Ferens, F. G. & Tarade, D. Hypoxia-inducible factor underlies von Hippel-Lindau disease stigmata. Elife 11 (2022). https://doi.org/10.7554/eLife.80774

(13) Wind, J. J. & Lonser, R. R. Management of von Hippel-Lindau disease-associated CNS lesions. Expert Rev Neurother 11, 1433-1441 (2011). https://doi.org/10.1586/ern.11.124

(14) Lonser, R. R. et al. von Hippel-Lindau disease. Lancet 361, 2059-2067 (2003). https://doi.org/10.1016/S0140-6736(03)13643-4

(15) Kaelin, W. G. Molecular basis of the VHL hereditary cancer syndrome. Nat Rev Cancer 2, 673-682 (2002).

(16) Costa, S. W. et al. A GABAergic and peptidergic sleep neuron as a locomotion stop neuron with compartmentalized Ca2+ dynamics. Nature Communications 10 (2019). https://doi.org/10.1038/s41467-019-12098-5

(17) Wu, Y., Masurat, F., Preis, J. & Bringmann, H. Sleep Counteracts Aging Phenotypes to Survive Starvation-Induced Developmental Arrest in C. elegans. Curr Biol 28, 3610-3624.e3618 (2018). https://doi.org/10.1016/j.cub.2018.10.009

(18) Busack, I. & Bringmann, H. A sleep-active neuron can promote survival while sleep behavior is disturbed. PLOS Genetics 19, e1010665 (2023). https://doi.org/10.1371/journal.pgen.1010665

(19) Powell-Coffman, J. A. & Coffman, C. R. Apoptosis: Lack of oxygen aids cell survival. Nature 465, 554-555 (2010). https://doi.org/10.1038/465554a

(20) Kruempel, J. C. P. et al. Hypoxic response regulators RHY-1 and EGL-9/PHD promote longevity through a VHL-1-independent transcriptional response. Geroscience 42, 1621-1633 (2020). https://doi.org/10.1007/s11357-020-00194-0

(21) Bach, M., Bergs, A., Mulcahy, B., Zhen, M. & Gottschalk, A. (2023).

(22) Leiser, S. F. et al. Cell nonautonomous activation of flavin-containing monooxygenase promotes longevity and health span. Science (2015). https://doi.org/10.1126/science.aac9257

Version 2

Abstract

A coordinated response to stress is crucial for promoting the short- and long-term health of an organism. The perception of stress, frequently through the nervous system, can lead to physiological changes that are fundamental to maintaining homeostasis. Activating the response to low oxygen, or hypoxia, extends healthspan and lifespan in C. elegans. However, despite some positive impacts, negative effects of the hypoxic response in specific tissues prevent translation of their benefits in mammals. Thus, it is imperative to identify which components of this response promote longevity. Here, we interrogate the cell-nonautonomous signaling pathway downstream of genetic activation of the hypoxic response. We find that HIF-1-mediated signaling in ADF serotonergic neurons is both necessary and sufficient for lifespan extension. Signaling through the serotonin receptor SER-7 in the GABAergic RIS interneurons is necessary in this process. Our findings also highlight the involvement of additional neural signaling molecules, including the neurotransmitters tyramine and GABA, and the neuropeptide NLP-17, in mediating longevity effects. Finally, we demonstrate that oxygen- and carbon-dioxide-sensing neurons act downstream of HIF-1 in this circuit. Together, these insights develop a circuit for how genetic induction of the hypoxic response cell-nonautonomously modulates aging and suggests valuable targets for modulating aging in mammals.

Introduction

The global population is aging at an unprecedented rate, posing significant challenges for healthcare systems worldwide1. The increasing burden of age-related diseases such as cardiovascular disease, diabetes, and neurodegeneration2 highlights the pressing need to improve healthspan, defined as the disease-free period of life3. Work over the past several decades has led to a better understanding of the conserved molecular and cellular mechanisms that regulate lifespan. One promising avenue has been the identification of environmental stressors that, when applied in a controlled manner, activate stress-response pathways to promote longevity. These interventions, including calorie restriction, exercise, hypoxia, heat shock, and cold exposure, extend lifespan and healthspan4-7. These stress-response pathways induce hormesis, a phenomenon in which a low dose of a stressor results in adaptive beneficial effects on cellular function8,9. Hormesis and the underlying stress-responses that promote longevity converge on cellular pathways, including oxidative stress, insulin signaling, autophagy, and protein homeostasis10, to promote longevity.

Multicellular organisms secrete a highly conserved set of neuromodulators including neurotransmitter and neuropeptide signals from the central nervous system. These signals integrate information about the organism’s external environment and internal state, and coordinate a physiological response in peripheral tissues. Within the context of aging, bioamine neurotransmitters contribute to the activation of many well-studied stress response mechanisms that promote longevity. For example, serotonin signaling modulates the mitochondrial unfolded protein response in C. elegans11, and regulates how food perception contributes to dietary restriction-mediated longevity in worms12 and flies13. Dopamine also modifies the longevity benefits of dietary restriction in invertebrates12,14, and polymorphisms in the dopamine D4 receptor are associated with longevity in humans15. Finally, adrenaline and its invertebrate analog, tyramine, have been linked to changes in the aging rate in both nematodes14,16-18 and mammals19. Hormone signals such as insulin20-22, growth factors23,24, and GnRH14,25 also contribute to longevity pathways across taxa. This growing body of work has led to great interest in manipulating longevity-promoting neural circuits to improve health26. While significant research effort focuses on pathways like autophagy, insulin signaling, and proteostasis in the context of aging, the role of hypoxia remains relatively underexplored. Currently, no longevity-promoting agents have been reported to target the hypoxic response.

The hypoxic response is highly conserved across species27-29, highlighting the relevance of this pathway in other organisms, including humans. In C. elegans, genetically activating the hypoxic response by knocking down or mutating the E3 ubiquitin ligase von hippel-lindau-1 (VHL-1)30 protein or stabilizing the hypoxia-inducible factor-1 (HIF-1) transcription factor6,31 is sufficient to extend lifespan. All of these interventions—environmental hypoxia, VHL-1 knockdown, and HIF-1 stabilization—require the intestinal enzyme flavin-containing monooxygenase-2 (fmo-2) to extend lifespan in C. elegans31. Hypoxic conditions also extend lifespan in a short-lived progeria mouse model32, and epidemiological studies indicate a correlation between hypoxia exposure and longevity33,34. However, the physiological changes induced by the hypoxic response are broad and involve adaptations such as increased vascularization, metabolic rewiring, and changes in cell survival pathways. In mammals, some of these same adaptations can be detrimental, as mutations in components of the hypoxic response have been linked to conditions like cancer and cardiovascular disease35-38. Additionally, HIF activity is essential to promote some forms of stress resistance to infection39,40 and oxidative stressors41,42 but can have detrimental effects on proteostasis43. This presents a key challenge: while hypoxia-induced longevity in a post-mitotic invertebrate model like C. elegans is promising, some of the mechanisms involved may not be directly translatable to mammals without causing deleterious side effects. Therefore, a mechanistic understanding of the individual cells, neural circuits, and pathways that mediate the beneficial but not detrimental effects of hypoxia on aging is essential to determine whether this circuit could be leveraged to improve human health.

Our previous work in C. elegans identified that stabilization of HIF-1 in neurons is sufficient to extend lifespan through the serotonin receptor, SER-7. This pathway eventually leads to the induction of fmo-2, a longevity gene expressed in the intestine31,44. In this study, we uncover key neural components of the longevity circuit initiated by genetic induction of the hypoxic response. Within this circuit, we identify individual cells, signals, and receptors necessary and/or sufficient to extend lifespan downstream of genetic activation of the hypoxic response. More specifically, we find serotonin signaling in the ADF serotonergic neurons is both necessary and sufficient to extend lifespan through genetic activation of the hypoxic response. This pathway signals through the serotonin receptor SER-7 in the RIS interneuron. We further demonstrate additional neurotransmitters (GABA and tyramine), and a neuropeptide (NLP-17) are critical for mediating these longevity effects. Finally, we identify that oxygen sensing neurons (URX, AQR, PQR and BAG) act downstream of neuronal HIF-1 in this circuit. Our insights into this longevity pathway provide a mechanistic understanding of how genetic activation of the hypoxic response delays aging and improves health.

Results

Serotonin signaling through the ADF neuron and the SER-7 receptor are necessary and sufficient for genetic activation of the hypoxic response to extend lifespan

Induction of the hypoxic response by targeted genetic manipulations can increase lifespan in C. elegans. These manipulations include decreasing activity of the VHL-1 E3 ubiquitin ligase that targets the transcription factor HIF-1 for degradation under hypoxia30 or by a mutation that stabilizes HIF-1 (HIF-1P621A)6,31 (Fig. 1A). We previously found that stabilizing HIF-1 in serotonergic neurons is sufficient to extend lifespan31. To determine which serotonergic neurons initiate HIF-1-mediated longevity, we generated strains with a nondegradable HIF-1 variant (HIF-1P621A)31,45 expressed under promoters specific to each of C. elegans’ three primary serotonergic neuron types—the ADF, NSM, and HSN neurons46. These transgenic worms were then crossed into the hif-1 null background to ensure that HIF stabilization in one neuron type could not feedback to modify HIF-1 activity in other cells. When we measured the lifespan of each strain relative to WT and hif-1 knockout controls, we observed that stabilizing HIF-1 in the ADF or NSM neurons significantly extended lifespan by 26% and 23%, respectively (Fig. 1B-C). HIF-1 stabilization in the HSN neurons had a smaller effect but significantly extended lifespan by 9% (Fig. 1D). This result indicates that modifying signaling in any serotonergic neuron is sufficient to induce some level of hypoxic response-mediated longevity. These data suggest that serotonergic neurons can partially substitute for each other to promote longevity in this pathway. In summary, the ADF and/or NSM serotonergic neurons play an essential role in longevity following genetic activation of the hypoxic response.

ADF serotonergic neurons are necessary and sufficient to extend lifespan downstream of genetic activation of the hypoxic response.

(A) Diagram of the conserved hypoxic response and genetic approaches to activate it, vhl-1 knockdown and HIF-1 stabilization. (B-D) Survival curves of WT, hif-1(ia4) knockout, hif-1(ia4); ADF::HIF-1S (B), hif-1(ia4); NSM::HIF-1S (C), and hif-1(ia4); HSN::HIF-1S (D) worms. N ≥ 256 (B), N ≥ 188 (C), and N ≥ 188 (D) worms per condition. (E) Survival curves of WT and ADF:tph-1 knockout worms on empty vector (EV) or vhl-1 RNAi. N ≥ 143 worms per condition. (F) Survival curves of WT and NSM:tph-1 knockout worms on empty vector (EV) or vhl-1 RNAi. N ≥ 193 worms per condition. (G) Survival curves of WT and tph-1(mg280); ADF:tph-1 rescue worms on empty vector (EV) or vhl-1 RNAi. N ≥ 178 worms per condition. (H) Survival curves of WT and tph-1(mg280); NSM:tph-1 rescue worms on empty vector (EV) or vhl-1 RNAi. N ≥ 160 worms per condition. (I) Gene expression of fmo-2 in WT, hif-1(ia4) knockout, hif-1(ia4); ADF:HIF-1S, and hif-1(ia4); NSM:HIF-1S worms. Significance in panels B-H is from a log-rank test comparing median survival. Significance in panel I is from a Student’s t-test 47. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001.

After identifying this role for ADF and NSM HIF-1 stabilization, we also asked whether genetically activating the hypoxic response also affected healthspan in three different measures of C. elegans mobility: pumping rate, thrashing rate, and movement speed. We found that both ADF- and NSM-specific HIF-1 stabilization had no effect on pumping rate in young (day 1 of adulthood) worms. In aged animals (day 12 of adulthood), however, NSM-, but not ADF-, specific HIF-1 stabilization rescued the pumping rate decline in hif-1 knockout compared to WT worms (Fig. S1C). Similarly, NSM-, but not ADF-, specific HIF-1 stabilization rescued the thrashing rate decline in the hif-1 knockout young and aged worms (Fig. S1D). ADF and NSM HIF-1 stabilization also had no effect on average or maximum movement speed at days 1 and 5 of adulthood (Fig. S1E-F). Together, these results indicate that genetic activation of the hypoxic response in NSM neurons but not in the ADF neurons could improve healthspan.

HIF-1 activity also interacts with multiple stress responses, including oxidative stress, proteotoxic stress, and infection41,42 39,40,43. To determine whether various components of this pathway affect both longevity and the response to other stressors, and given that serotonin signaling modulates the mitochondrial unfolded protein response (mt-UPR)48, we asked whether stabilizing HIF-1 in these serotonergic neurons may also affect the mt-UPR. We found that HIF-1 stabilization in either the ADF or the NSM decreased the expression of hsp-6 (Fig. S1G), a downstream effector of the mt-UPR49. This could suggest either that the mt-UPR response is impaired in these worms, or that HIF-1 stabilization decreases proteotoxic stress leading to a lower basal level of hsp-6.

To further test the necessity of the most impactful (for lifespan) serotonergic neurons in genetic activation of the hypoxic response-mediated longevity, we used transgenic animals in which tph-1, the rate-limiting enzyme in serotonin synthesis, was knocked out exclusively in the ADF or the NSM serotonergic neurons12,50. We then measured the lifespans of these strains on empty vector (EV) and vhl-1 RNAi to determine the necessity of ADF and NSM serotonergic signaling in vhl-1-mediated longevity. Successful RNAi knockdown was confirmed with qPCR validation (Fig. S1A). We found that serotonin synthesis in ADF but not NSM neurons was required for vhl-1 knockdown to extend lifespan (Fig. 1E-F). This indicates that while HIF stabilization in multiple serotonergic neurons can extend lifespan, ADF serotonin signaling is required for lifespan extension in response to vhl-1 knockdown.

To test if serotonin production in a single neuron type is sufficient for genetic activation of the hypoxic response to extend lifespan, we rescued tph-1 expression in a tph-1 null background using promoters specific to the ADF or the NSM neurons12. The lifespans of these strains on EV or vhl-1 RNAi were then measured. As expected and consistent with our previous publication31, the lifespan of the tph-1 knockout strain was not extended by vhl-1 knockdown (Fig. S1B). Consistent with the ADF neuron’s requirement for vhl-1-mediated longevity (Fig. 1E-F), serotonin synthesis in the ADF neurons was sufficient for vhl-1 knockdown to extend lifespan while NSM serotonin synthesis was not (Fig. 1G-H). The enzyme fmo-2 is induced in the intestine under hypoxia and is required for lifespan extension by environmental hypoxia, HIF-1 stabilization, and vhl-1 depletion31. To test whether ADF- or NSM-specific HIF-1 stabilization also induces fmo-2, we measured fmo-2 transcription in this strain using qPCR. We found that ADF and NSM HIF-1 stabilization significantly induced fmo-2 expression relative to WT controls (Fig. 1I), suggesting a model in which ADF signaling extends lifespan through the same mechanism as vhl-1 knockdown and environmental hypoxia. Together, these data indicate that while HIF-1 stabilization in any serotonergic neuron is sufficient to extend lifespan, only ADF serotonin production is necessary and sufficient for vhl-1-mediated longevity. This suggests that under normal physiological conditions the serotonergic ADF neurons propagate a signal in response to whole body stabilized HIF-1.

After identifying ADF neurons as the most central serotonergic neurons in longevity following genetic activation of the hypoxic response, we next sought to determine the downstream receptor responding to serotonin signaling from ADF neurons. C. elegans have six known serotonin receptors—SER-1, SER-4, SER-5, SER-7, MOD-1, and LGC-50. Of these six receptors, our previous findings showed that ser-7 expression is required for vhl-1 knockdown to induce fmo-2 and extend lifespan31. SER-7 is a G protein-coupled receptor (GPCR) with high sequence identity to the 5-HT7 receptor in mammals51,52. ser-7 is thought to be expressed in 27 of C. elegans’ 302 neurons53 and contributes to many aspects of physiology such as reproduction and pharyngeal pumping51. Consequently, broad manipulation of SER-7 signaling is not an ideal method for exclusively modifying aging. To identify more specific targets for lifespan extension within this pathway, we asked whether ser-7 expression in any subset of neurons is sufficient to rescue the vhl-1 mediated fmo-2 induction in a ser-7 null background.

We created transgenic strains with ser-7 expression under the control of 11 different cell-specific promoters in a ser-7 null background. These promoters were selected to rescue ser-7 expression in the following neuronal populations: whole-body rescue (ser-7p::ser-7), interneuron rescue (glr-1p::ser-7), bioaminergic neuron rescue (cat-1p::ser-7), glutamatergic neuron rescue (eat-4p::ser-7), GABAergic neuron rescue (unc-47p::ser-7), cholinergic neuron rescue (unc-17p::ser-7), sensory neuron rescue (osm-6p::ser-7), GABAergic motor neuron rescue (unc-25p::ser-7), cholinergic motor neuron rescue (acr-2p::ser-7), M3 and M4 neuron rescue (ceh-28p::ser-7), and intestinal rescue (vha-6p::ser-7) in a ser-7 null background. To measure whether ser-7 expression restores vhl-1 mediated longevity, we injected each construct into ser-7 null worms crossed with a single-copy transcriptional reporter for fmo-2 (fmo-2p::mCherry). Because fmo-2 is induced by and required for hypoxia-mediated longevity31, we utilized its induction as an efficient screening tool to identify components of this longevity pathway.

We first confirmed that ser-7 knockout attenuates vhl-1-mediated fmo-2 induction (Fig. 2A). The whole-body rescue of ser-7 expression under the endogenous ser-7 promoter restored fmo-2 induction to 91% of WT controls (Fig. 2A). Of the 10 cell-specific ser-7 rescues, expression in interneurons, bioaminergic neurons, glutamatergic neurons, or GABAergic neurons fully rescued vhl-1-mediated fmo-2 induction to ≥ 100% of the WT control response (Fig. 2A, pink bars). ser-7 rescue in GABAergic motor neurons or in sensory neurons showed a partial rescue (defined as a 50-99% increase from fmo-2 induction in the ser-7 knockout) (Fig. 2A, orange bars). In contrast, rescuing ser-7 in cholinergic motor neurons, cholinergic neurons, M3 and M4 neurons, or in the intestine did not rescue (Fig. 2A, gray bars). Taken together, 6 of the 10 cell-specific ser-7 rescues at least partially restored vhl-1-mediated fmo-2 induction, including some that have no known overlap in ser-7 expression patterns. Collectively, these data suggest that serotonin receptor ser-7 expression in multiple neurons may be sufficient to convey the hypoxic signal to the intestine and induce fmo-2.

ser-7 expression in the GABAergic RIS neuron is required for genetic activation of the hypoxic response to extend lifespan.

(A) Quantification of fmo-2p::mCherry (WT positive control), fmo-2p::mCherry; ser-7(tm1325) knockout (negative control), and fmo-2p::mCherry; ser-7(tm1325) knockouts with ser-7 rescued in the whole body (ser-7p::ser-7), interneurons (glr-1p::ser-7), bioaminergic neurons (cat-1p::ser-7), glutamatergic neurons (eat-4p::ser-7), GABAergic neurons (unc-47p::ser-7), GABAergic motor neurons (unc-25p::ser-7), sensory neurons (osm-6p::ser-7), cholinergic motor neurons (acr-2p::ser-7), cholinergic neurons (unc-17p::ser-7), M3 & M4 neurons (ceh-28p::ser-7), and the intestine (vha-6p::ser-7) on empty vector (EV, WT control shown) or vhl-1 RNAi (all strains shown). Bar height indicates the mean fluorescence on vhl-1 RNAi of each genotype normalized to the empty vector control value from that genotype. The dashed line indicates the vhl-1-mediated fmo-2 induction of WT positive control. N ≥ 45 worms per condition. Error bars indicate SEM. Significance is from a two-way ANOVA (fmo-2 induction ∼ genotype*RNAi) and post-hoc Tukey HSD test47. Stars signify results from this post-hoc test comparing the fmo-2 induction of each strain on EV RNAi the fmo-2 induction of that strain on vhl-1 RNAi. (B) A table showing the 10 ser-7 expressing candidate neurons and their expression patterns among the ser-7 cell-specific strains that successfully rescued fmo-2 induction in Fig. 2A. Pink rows indicate rescues that full restored vhl-1 mediated fmo-2 induction, orange rows indicate rescues that partially restored vhl-1-mediated fmo-2 induction, and gray rows signify unsuccessful rescue constructs from the data in Fig. 2A. (C) Survival curve of WT and RIS ablation (Ex[srsx-18p::caspase-3(p12)-nz]; Ex[srsx-18p::cz-caspase-3 (p17)]; srsx-18p::GFP) worms on empty vector (EV) and vhl-1 RNAi. N ≥ 192 worms per condition. (D) Survival curve of WT and ser-7 (tm1325); ser-7 rescue in the RIS neuron (flp-11p::ser-7) worms on EV and vhl-1 RNAi. N ≥ 144 worms per condition. Significance in panels C-D is from a log-rank test comparing median survival. In all panels, NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

We compiled a table containing the top candidates for the ser-7 neuron based on which rescue constructs restored ser-7 expression (simplified table in Fig. 2B, full table in Fig. S2A). We hypothesized that neurons present in many of the successful rescue constructs would be most important in the vhl-1-mediated longevity circuit. Conversely, we hypothesized that candidate neurons within rescues that did not restore a WT-like response are less likely to be important (Fig. 2B; Fig. S2A). We acknowledge that this analysis strategy is biased toward neurons rescued in a greater proportion of the transgenic strains, which could lead to false negatives. From this analysis, we identified 10 neurons that were rescued in 3-4 of the 6 successful rescue constructs (Fig. 2B). Five of these neurons (DVB, FLP, M1, RME_LR, and RME_DV) were also present in unsuccessful rescue constructs, so they were not considered for follow-up experiments. Of the remaining five neurons (RIS, AUA, AVC, LUA, and VD_DD), the top candidate was the RIS neuron, because RIS was present in the highest number (4 as shown in Fig. 2B) of successful rescue constructs.

To validate whether the RIS neuron is a key signaling cell in vhl-1-mediated longevity, we first genetically ablated the RIS neuron to test its necessity. Our results showed that vhl-1 RNAi does not extend lifespan of RIS ablated worms, indicating that the RIS neuron is completely required for vhl-1 knockdown to extend lifespan (Fig. 2C). To test the sufficiency of RIS signaling, we next rescued ser-7 expression in the RIS neuron in the ser-7 null background under the flp-11 promoter. flp-11 is expressed in the RIS neuron54 and uv1 neuroendocrine cells55 that do not endogenously express ser-753. We first confirmed that ser-7 knockout worms are partially required for vhl-1 RNAi to extend lifespan (Fig. S2B, Cox Regression, p = 0.003, **). Significantly, ser-7 rescue in the RIS neuron (ser-7; flp-11p::ser-7) restores the lifespan extension by vhl-1 RNAi to the same degree as WT worms (Fig. 2D). These data suggest that ser-7 expression in the RIS neuron only is sufficient for vhl-1-mediated longevity. However, one limitation of this approach is that non-physiological expression of ser-7 in the uv1 cells could play a role in restoration of WT-like vhl-1 mediated fmo-2 induction. Together, these data are consistent with a model where genetic activation of the hypoxic response stabilizes HIF-1 in the ADF neurons, which modifies serotonin signaling to the SER-7 receptor on the RIS interneuron.

Tyraminergic signaling from the RIM neuron and the tyramine receptor TYRA-3 are required for vhl-1 knockdown to extend lifespan

After identifying neuron subtypes involved in serotonin signaling downstream of genetic activation of the hypoxic response, we next set out to identify whether other neurotransmitters act in the pathway. To answer this question, we obtained mutants deficient in the production of each type of neurotransmitter in C. elegans (Fig. 3A) and measured their lifespan on control and vhl-1 RNAi. Out of the six neurotransmitters we tested, blocking the production of GABA (unc-25, Fig. 3B) or tyramine + octopamine (tdc-1, Fig. 3C) attenuated longevity on vhl-1 RNAi. The requirement of GABA is consistent with our finding that ser-7 expression on the GABAergic RIS neuron is also necessary for vhl-1-mediated longevity (Fig. 2), connecting the serotonergic and GABAergic components of this circuit.

GABA and tyramine synthesis are required for vhl-1 knockdown to extend lifespan.

(A) Summary table of the necessity of each C. elegans neurotransmitter for vhl-1 RNAi-mediated longevity. (B-E) Survival curves of WT and unc-25(e156) (GABA deficiency) (B), tdc-1(n3419) (tyramine + octopamine deficiency) (C), tbh-1(n3247) (octopamine deficiency) (D), and RIC ablation (Ex[tbh-1p::caspase-3(p12)-nz]; Ex[tbh-1p::cz-caspase-3 (p17)]; tbh-1p::GFP) (E) worms on empty vector (EV) and vhl-1 RNAi. N ≥ 310 (B), ≥ 201 (C), ≥ 258 (D), and ≥ 258 (E) worms per condition. Significance in panels B-E is from a log-rank test comparing median survival. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

Tyramine (analogous to mammalian adrenaline56) is synthesized from tyrosine by tyrosine decarboxylase (tdc-1) and can be further converted to octopamine (analogous to mammalian noradrenaline) by tyramine β-hydroxylase (tbh-1) in C. elegans57. Therefore, the tdc-1 knockout is deficient in both tyramine and octopamine synthesis (Fig. 3C) and the tbh-1 knockout (Fig. 3D) is only deficient in octopamine synthesis. Since worms lacking only octopamine (tbh-1, Fig. 3D) were long-lived on vhl-1 RNAi but worms lacking both tyramine and octopamine (tdc-1, Fig. 3C) were not, we can conclude that tyramine is required for genetic activation of the hypoxic response to extend lifespan. To further test whether tyramine but not octopamine is involved in this pathway, we generated a strain in which the primary octopamine-producing neuron, RIC57, is genetically ablated. We found that the RIC neuron is not required for vhl-1-mediated longevity (Fig. 3E), further supporting that tyramine but not octopamine is a key neurotransmitter in the circuit driving longevity after genetic activation of the hypoxic response.

In addition to octopamine (tbh-1, Fig. S3C), we also observed that acetylcholine (unc-17, Fig. S3A) and glutamate (eat-4, Fig. S3B) were not required for lifespan extension by vhl-1 RNAi. Our lifespans of the dopamine synthesis mutant cat-2 indicated that dopamine is inconsistently required for genetic activation of the hypoxic response to extend lifespan. In three biological replicates, we observed a full requirement, partial requirement, and no requirement for dopamine synthesis in vhl-1 RNAi-mediated longevity. When analyzed together, the net result suggests that dopamine is partially required for vhl-1-mediated longevity (Fig. S3C, Cox Regression for interaction between genotype and vhl-1 RNAi, p < 0.0001, ****). We did not follow up on dopamine due to the inconsistency. Together, the results from this neurotransmitter screen support the conclusion that in addition to serotonin, both GABA and tyramine play a role in vhl-1-mediated longevity.

Nematodes synthesize GABA in 26 of 302 neurons while tyramine is only synthesized in two cells58. As a result, identifying highly specific components of tyraminergic signaling is more feasible than mapping GABAergic signaling components. Therefore, we sought to determine the tyramine-producing neuron(s) and tyramine receptor(s) that contribute to vhl-1-mediated longevity. In C. elegans, the two canonically tyraminergic cell types (the RIM neuron and the uv1 neuroendocrine cells) express tdc-1 but not tbh-158. To determine which tyraminergic cell(s) act in thevhl-1-mediated longevity pathway, we expressed tdc-1 under promoters specific to either RIM or uv1 in a tdc-1 null background14. We then measured whether the lifespans of these tyramine rescue strains could be extended by vhl-1 RNAi. Our results showed that vhl-1 RNAi extended lifespan in the RIM rescue strain (Fig. 4A), but not in the uv1 rescue (Fig. 4B). This finding demonstrates that tyramine synthesis in the RIM neuron is sufficient for vhl-1 RNAi to extend lifespan. Notably, both tyramine rescue constructs lived longer than WT controls (Fig. 4A-B). This could indicate that altering tyramine signaling modifies lifespan both within and independently of the vhl-1-mediated longevity pathway.

The RIM neuron and the tyramine receptor tyra-3 act in the vhl-1-mediated longevity pathway.

(A-B) Survival curves of WT, RIM rescue (tdc-1 (n3419); Ex[ocr-4p::tdc-1]) (A), and uv1 tdc-1 rescue (tdc-1 (n3419); Ex[gcy-13p::tdc-1]) (B) strains on empty vector (EV) and vhl-1 RNAi. N ≥ 126 (A) and N ≥ 140 (B) worms per condition. (C) Survival curves of TU3311(unc-119p::sid-1) and vhl-1(ok161); TU3311(unc-119p::sid-1) worms on empty vector (EV) and tyra-3 RNAi. N ≥ 215 worms per condition. Significance in all panels is from a log-rank test comparing median survival. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

To identify the tyramine receptor(s) necessary for genetic activation of the hypoxic response to extend lifespan we used RNAi to knockdown each of C. elegans’ four known tyramine receptors59-62 (Fig. 4C). Since tyramine receptors are mostly expressed in neurons, and RNAi uptake into some neurons is less efficient than other cell types, we used a strain with enhanced neuronal RNAi uptake (TU3311, unc-119p::sid-1) crossed into vhl-1 knockout worms. This neural enhanced RNAi uptake strain was not used previously in this work for vhl-1 RNAi experiments because vhl-1 RNAi knockdown extends lifespan in WT worms31. The generated vhl-1; unc-119p::sid-1 strain was long-lived compared to the control strain unc-119p::sid-1 (Fig. 4C). Successful RNAi knockdown was validated via qPCR (Fig. S4A). Our lifespan results show that the tyramine receptor tyra-3 was fully required for vhl-1-mediated longevity (Fig. 4C), but the tyramine receptors tyra-2 and ser-2 were not (Fig. S4B-C). Another tyramine receptor, Igc-55, showed a partial requirement for vhl-1-mediated longevity (Fig. S4D). Although there was a minor lifespan extension by vhl-1 knockout on Igc-55 RNAi (Fig. S4D), we did observe a significant interaction between lgc-55 knockdown and genotype on lifespan (Cox Regression, p < 0.0001, ****), indicating a partial requirement. In summary, these data indicate that tyramine signaling from the RIM neuron to the TYRA-3 receptor is required for vhl-1-mediated longevity, while the LGC-55 receptor may play a partial role. Interestingly, TYRA-3 is highly expressed on the low oxygen/high carbon dioxide-sensing BAG neuron as well as in the intestine53,63, the site of fmo-2 induction during hypoxia31.

Oxygen and carbon dioxide sensing neurons act downstream of serotonin signaling

Given that tyra-3 is highly expressed in the canonical low-oxygen sensing neuron BAG53, we next explored whether neurons responsible for sensing high- and low-oxygen conditions are essential for genetic activation of the hypoxic response to extend lifespan. C. elegans have four oxygen sensory neurons: the URX, PQR, and AQR neurons detect high levels of oxygen63, the BAG neuron responds to low levels of oxygen64 and high levels of carbon dioxide65 (Fig. 5A). We hypothesized that these oxygen sensing cells may perceive hypoxic conditions and help initiate the hypoxic response. Alternatively, these oxygen sensing neurons could modify behavior in response to hypoxia but not play a role in longevity. We found that vhl-1 knockdown was unable to extend lifespan when we genetically ablated the three high-oxygen sensing neurons (URX, AQR, PQR, Fig. 5B). Interestingly, the low O2/high CO2 responsive BAG neurons were also required for vhl-1-mediated longevity (Fig. 5C). To test the epistasis of oxygen sensing neurons and serotonergic neurons in this circuit, we next crossed the URX/AQR/PQR ablation worms or BAG ablation worms into C. elegans with HIF-1 stabilized in the ADF neurons. Ablation of either the URX/PQR/AQR or the BAG neurons abrogated lifespan extension in the ADF HIF-1 stabilized worms. This suggests that oxygen sensing neurons act downstream of ADF HIF-1 stabilization in this pathway (Fig. 5D-E).

High- and low-oxygen sensing neurons are important for genetic activation of the hypoxic response to extend lifespan.

(A) Diagram of oxygen sensing neurons in C. elegans. (B) Survival curves of WT and URX/PQR/AQR ablated worms (qaIs2241 [gcy-36::egl-1 + gcy-35::GFP + lin-15(+)]) on empty vector (EV) and vhl-1 RNAi. N ≥ 279 worms per condition. (C) Survival curves of WT and BAG ablated worms (Ex[gcy-31p::caspase-3(p12)-nz]; Ex[gcy-31p::cz-caspase-3 (p17)]; gcy-31p::GFP) on empty vector (EV) and vhl-1 RNAi. N ≥ 188 worms per condition. (D) Survival curves of WT, hif-1(ia4); ADF-HIF-1S, URX/PQR/AQR (qaIs2241 [gcy-36::egl-1 + gcy-35::GFP + lin-15(+)]) ablation, and URX/PQR/AQR (qaIs2241 [gcy-36::egl-1 + gcy-35::GFP + lin-15(+)]) ablation; hif-1(ia4); ADF-HIF-1S worms. N ≥ 150 worms per condition. (E) Survival curves of WT, hif-1(ia4); ADF-HIF-1S, BAG (Ex[gcy-31p::caspase-3(p12)-nz]; Ex[gcy-31p::cz-caspase-3 (p17)]; gcy-31p::GFP) ablation, and BAG (Ex[gcy-31p::caspase-3(p12)-nz]; Ex[gcy-31p::cz-caspase-3 (p17)]; gcy-31p::GFP) ablation; hif-1(ia4); ADF-HIF-1S worms. N ≥ 186 worms per condition. Significance in all panels is from a log-rank test comparing median survival. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

The neuropeptide NLP-17 and its receptors contribute to vhl-1-mediated longevity

After identifying neurons, neurotransmitters, and neuroreceptors acting in the vhl-1-mediated longevity circuit, we wondered how this neural circuit ultimately propagates information about hypoxic conditions to the intestine and induces the pro-longevity gene fmo-2. The C. elegans nervous system does not directly innervate the intestine. Instead, neurosignaling molecules bind to receptors on peripheral tissues through packaging in dense-core vesicles and subsequent release into and diffusion through pseudocoelomic fluid66,67. These dense-core vesicles are packaged with a wide variety of neuropeptide signals, although there is also evidence that bioaminergic neurotransmitters like serotonin, dopamine, and adrenaline/noradrenaline can also signal through this dense-core vesicle mechanism67-69.

We first tested if inducing the hypoxic response through vhl-1 RNAi extends lifespan in a mutant strain lacking the dense-core vesicle packaging gene unc-31. We observed that unc-31 was required for vhl-1 knockdown to extend lifespan (Fig. 6A). To examine whether neuropeptide signals carried by dense-core vesicles are involved, we knocked down two neuropeptide processing enzymes—egl-3 and egl-2170. Successful knockdown was validated with qPCR (Fig. S5A). Interestingly, egl-21 but not egl-3 was required for vhl-1-mediated longevity (Fig. 6B-C). This could be because egl-21 is the only enzyme known to perform the carboxypeptidase E cleavage step of neuropeptide processing71, while there are four known proprotein convertases with similar function to egl-370,72.

The neuropeptide nlp-17 and its receptors npr-37 and npr-43 are required for vhl-1-mediated longevity.

(A) Survival curves of unc-31(e169) and WT worms on empty vector (EV) and vhl-1 RNAi. N ≥ 242 worms per condition. (B-C) Survival curves of TU3311 (unc-119p::sid-1) and vhl-1(ok161); TU3311 (unc-119p::sid-1) worms on empty vector (EV) and egl-3 (B) or egl-21 (C) RNAi. N ≥ 209 worms (B) and N ≥ 185 worms (C) per condition. (D) Summary table of two sets of neuropeptide ligands/receptors that blunted vhl-1-mediated lifespan. (E-G) Survival curves of TU3311 (unc-119p::sid-1) and vhl-1(ok161); TU3311 (unc-119p::sid-1) worms on empty vector (EV) and npr-37 (E), npr-43 (F), or nlp-17 (G) RNAi. N ≥ 288 worms (E), N ≥ 309 worms (F), and N ≥ 309 worms (G) per condition. Significance in panels A-C and E-G are from log-rank test comparing median survival. Cox Regression for a significant interaction between npr-43 knockdown and vhl-1 knockout on lifespan, p <0.01, **. Cox regression for a significant interaction between nlp-17 knockdown and vhl-1 knockout on lifespan, p < 0.05, *. NS. = p > 0.05, * = p < 0.05, ** = p < 0.01, *** = p < 0.001, and **** = p < 0.0001. In all panels, three replicates were plotted together, and all statistics include a Bonferroni correction for multiple comparisons.

We next tested a group of 6 genes corresponding to neuropeptide/receptor binding pairs that blunted vhl-1-mediated fmo-2 induction (Fig. 6D). These genes, along with their corresponding receptors, were targeted for RNAi and tested for lifespan in TU3311 (unc-119p::sid-1) and TU3311 (unc-119p::sid-1); vhl-1 (ok161) strains. We found that the npr-3 receptor and its ligands, flp-15 and flp-21 were not required for vhl-1-mediated longevity. (Fig. S5B-D). However, knocking down the neuropeptide nlp-17 partially blocked vhl-1-mediated longevity (Fig. 6E, Cox regression p < 0.05). Consistently, knocking down nlp-17’s receptors also completely (npr-37, Fig. 6F) or partially (npr-43, Fig. 6G, Cox regression p < 0.001) blocked vhl-1-mediated longevity. Successful knockdown of npr-37 and npr-43 was validated with qPCR (Fig. S5A). The nlp-17 ligand result was validated using a genetic knockout (nlp-17 (ok3461)), which completely prevented vhl-1 knockdown from extending lifespan (Fig. S5E). To test whether nlp-17 acts downstream of other signals in this circuit, we measured the expression of nlp-17 in ADF- and NSM-specific HIF-1 stabilized strains. Compared to WT and hif-1 KO controls, we observed no change in nlp-17 expression when HIF-1 was stabilized in the ADF or NSM serotonergic neurons (Fig. S5F). This could suggest either that nlp-17 signaling acts in parallel to serotonergic signaling following genetic activation of the hypoxic response. Alternatively, neuronal HIF-1 stabilization may modify nlp-17 splicing or translation without resulting in detectable differences in mRNA levels. Together, these data indicate NLP-17 signaling is required for longevity following genetic activation of the hypoxic response, although whether this peptide is synthesized or released in response to hypoxic response remains unclear.

Taken together, these results support a model in which HIF-1 stabilization in the ADF serotonergic neurons leads to modified serotonin signaling from the ADF to the SER-7 expressing and GABA-producing RIS neuron (Fig. 7A-B). Oxygen sensing neurons act downstream of HIF-1 stabilization in the ADF neuron in this longevity circuit. Tyramine produced by the RIM neuron and neuropeptide (NLP-17) signaling are also required for genetic activation of the hypoxic response to extend lifespan, although it remains unclear whether these signals act upstream, downstream, or in parallel to the serotonergic and GABAergic components in this pathway.

Working model of longevity following genetic activation of the hypoxic response.

A) Schematic of working model. B) Table summarizing the necessity, sufficiency, and epistasis of neural signaling components of the working model.

Discussion

In this study, we interrogated a complex signaling pathway through which genetic induction of the hypoxic response extends lifespan in C. elegans. Within this pathway, we demonstrate that: 1) stabilization of HIF-1 in the ADF serotonergic neurons modifies signaling to the SER-7 receptor on the GABAergic and neuropeptidergic RIS neuron (Fig. 1-2); 2) tyraminergic signaling through the RIM neuron and the TYRA-3 receptor (Fig. 3-4) are required for vhl-1-mediated longevity; 3) oxygen and carbon-dioxide sensory neurons act downstream of HIF-1 stabilization and serotonergic ADF neurons to extend lifespan (Fig. 5); and 4) in addition to neurotransmitter signals, the neuropeptide NLP-17 and its receptors NPR-37 and NPR-43 are required for genetic activation of the hypoxic response to extend lifespan (Fig. 6). Ultimately, these neural signals converge on the intestine, to activate FMO-2 and extend lifespan (working model summarized in Fig. 7). Evidence for the necessity, sufficiency, and epistatic relationships between each signal are summarized in Fig. 7B. In brief, all signaling molecules presented in this work are necessary for vhl-1 to extend lifespan. Rescuing ADF serotonin production, RIS ser-7 expression, and RIM tyramine production is sufficient for vhl-1 to extend lifespan. The sufficiency of oxygen sensing neurons BAG and UPA/PQR/AQR, and the neuronal signals of GABA, NLP-17, and TYRA-3 to restore vhl-1 mediated longevity remains unclear. All signals act downstream of vhl-1. ADF HIF-1 stabilization and SER-7 signaling act upstream of fmo-2 induction, and the oxygen sensing neurons (BAG, UPA/PQR/AQR) act upstream of or in parallel to ADF HIF-1 stabilization.

The hypoxic response is highly conserved across species, highlighting the potential relevance of this pathway in other organisms, including humans. In nematodes, activating the hypoxic response promotes health and longevity. However, the physiological changes induced by the hypoxic response are broad and involve adaptations that can drive tumorigenesis in mammals. Therefore, a mechanistic understanding of the individual cells, signals, and circuits that mediate the beneficial effects of hypoxia is essential. In this work, the discovery that HIF-1 stabilization in the ADF neurons extends lifespan by 26% demonstrates the potential for targeted manipulations within this pathway to have large effects on lifespan. Additionally, our data identify a previously unknown role for tyramine/adrenaline in genetic activation of the hypoxic response as well as establish NLP-17, a neuropeptide with no previous known function, as a signal in vhl-1-mediated longevity. This work also demonstrates that the oxygen sensing AQR, PQR, URX, and BAG neurons act downstream of the serotonergic ADF neuron in genetic activation of the hypoxic response, indicating an interaction between internal and external oxygen sensing mechanisms in longevity. Together, the mapping of these cells, signaling molecules, and receptors acting downstream of genetic activation of the hypoxic response lays the foundation to discover druggable targets that selectively modulate aging in humans without negative side effects.

Other studies in C. elegans have also identified neural networks involving the hypoxic response. For example, previous work found a role for HIF-1 in hypoxia-mediated behaviors, such as food-dependent hyperoxia (excessive oxygen) avoidance73. Hyperoxia avoidance requires hif-1 expression both in the neurons and the tyraminergic uv1 neuroendocrine cells. Interestingly, serotonin signaling from the ADF neuron is also required for WT-like hyperoxia avoidance73. While this HIF-1-mediated behavioral circuit has parallels with the HIF-1-mediated longevity circuit interrogated in this work, these pathways do diverge: the key tyraminergic cell type implicated in hyperoxia avoidance is the uv1 gonadal neuroendocrine cells, while the RIM tyraminergic neurons played a role in longevity (Fig. 4). Together, the partial overlap between this behavioral and longevity circuit implies that modifying neural signaling to induce the hypoxic response and extend lifespan may also alter behavioral phenotypes. Future research in this area should interrogate exactly where these and any other HIF-1-mediated circuits overlap and diverge.

While this study identifies many neural signals required for vhl-1 knockdown or knockout to extend lifespan, one key limitation of this work is the potential differences between genetic and environmental methods of inducing the hypoxic response. While vhl-1 knockdown or knockout leads to HIF-1 stabilization by blocking its proteasomal degradation, it also results in hydroxylated but stable HIF-1. This contrasts with environmental hypoxia, in which HIF-1 remains stable because it cannot be hydroxylated. While HIF-1 is stabilized and localized to the nucleus in both cases, there are differences in transcriptional outcomes between stable hydroxylated and unhydroxylated states74,75. Additionally, we did not explore the effects of alternative genetic activators of the hypoxic response such as HIF-1 hydroxylase PHD/EGL mutants, which may provide further insight into how different manipulations of the hypoxic response impact longevity. Future work should interrogate the similarities and differences between the circuits driving longevity in response to environmental hypoxia, vhl-1 knockdown, HIF-1 stabilization, and PHD/EGL knockdown.

The circuit-mapping approaches employed in this work are also impacted by limitations in cell-specific genetic modifications and in the use of RNAi knockdown. For example, cell-specific rescue constructs can sometimes lead to unintended rescues in other cell types due to cell-nonautonomous signaling. Because all serotonin-producing neurons also express the serotonin reuptake transporter mod-5, serotonin produced by one cell in our rescue strains could be taken up by other serotonin-producing neurons, leading to unintended signaling effects. This may also be true of the uv1 and RIM tyraminergic rescue strains, although little is known about tyramine reuptake in C. elegans. Two cells identified via tissue-specific rescue experiments (ADF and RIS, Fig. 1G and Fig. 2D) were also found to be necessary via cell-specific knockout (ADF tph-1 KO, Fig. 1E; RIS ablation, Fig. 2C), decreasing the likelihood of a false positive from the rescue strain technique. However, the role of the RIM neuron was identified via a tdc-1 rescue strain and was not validated using a RIM-specific knockout (Fig. 7B). Therefore, the contribution of RIM signaling to this circuit is less well-validated, and a RIM-specific tdc-1 knockout strain should be examined in future work. Finally, additional genetic controls could better support the role of RIS-specific ser-7 expression in genetic activation of the hypoxic response. For example, a ser-7 expressing neuron that was not a hit in our screen could also be ablated and tested for necessity in vhl-1 mediated longevity. This experiment would test whether the ability of RIS ablation to attenuate vhl-1-mediated longevity is not a false positive driven by any disruption to ser-7 expression.

Additionally, although RNAi experiments were validated using sequence confirmation and qPCR, not all RNAi hits were corroborated with genetic knockouts. This leaves open the possibility that RNAi-induced knockdown effects differ from complete genetic ablation, or that production of a given RNAi may modify bacterial metabolism in a way that indirectly modifies the hypoxic response in C. elegans. Finally, while the use of RNAi knockdown and genetic knockouts establishes the necessity of many signals within the vhl-1-mediated longevity circuit, the exact directionality of these signals remains unclear. It is possible that increased, decreased, or pulsatile changes in signaling through these bioamines and neuropeptides are required for genetic activation of the hypoxic response to extend lifespan. Work on C. elegans reversal behavior has also revealed an antagonistic relationship between RIM and RIS activity facilitated by both chemical (neuropeptide and tyramine) and electrical (gap junction) signaling76,77. This known interaction should also be interrogated in the context of how these cells may communicate following genetic induction of the hypoxic response. Future work in this area could use tools to measure or modify neuronal activity, such as calcium imaging or optogenetics, to begin answering these questions.

While many individual neurosignaling components are essential for genetic activation of the hypoxic response to extend lifespan, their epistasis is unclear (Fig. 7B). Most components of the pathway identified in this work act downstream of vhl-1 depletion, and upstream of fmo-2 induction (summarized in Fig. 7A-B). However, the order of each signal between these two endpoints is only predicted based on C. elegans neural wiring and the overlap between various identified signals and cells. For example, we hypothesize in our working model that GABA may be produced by the RIS neuron in this circuit because RIS is the primary GABAergic neuron required for vhl-1-mediated longevity. Alternatively, it is possible that GABA is produced by a different cell that either acts in series or in parallel with RIS signaling. In order to determine the order of each signaling component, future studies should generate genetic manipulations to each signaling component that may mimic vhl-1 knockout to promote longevity, cross these new strains into knockouts of other signals required for vhl-1 mediated longevity; and measure the lifespans of each double and triple mutant. This approach would also narrow down which signals are downstream of the genetic activation of the hypoxic response, and which are sufficient to extend lifespan upstream of the hypoxic response in a normoxic environment. One notable target for further exploration is the SER-7 expressing RIS neuron, which plays a role in sleep77 and stress resistance78, and can extend lifespan when optogenetically activated under normoxic conditions79.

Another limitation of this work is that it uses lifespan as the main readout for organismal health. While genetic manipulations that extend lifespan often improve stress resistance and healthspan8,80, longevity manipulations can also have adverse effects on reproduction81,82 and behavior83,84. In this study, we find that HIF-1 stabilization in the ADF neurons does not prevent the deleterious effects of hif-1 knockout on mobility, but that HIF-1 stabilization in the NSM neurons may attenuate age-related decline in pumping and thrashing (Fig. S1). However, future work should examine whether other modifications to this pathway, such as manipulations to RIM, RIS, or NLP-17 signaling, influence healthspan in addition to lifespan. It will be important for future studies to determine whether various components of this pathway affect both longevity and the response to different types of stressors like oxidative stress, proteotoxic stress, and infection, as HIF-1 activity also interacts with multiple stress responses41,42 39,40,43.

While our understanding of this circuit is still incomplete, our findings point to several promising signaling components that could be targeted for longevity interventions. Manipulating serotonin and tyramine/adrenaline signaling pathways may offer strategies for extending lifespan, with minimal pleiotropic effects if precisely controlled. These neural targets have great potential for longevity therapeutics due to 1) the small number of neurons that can control aging-related pathways in the entire organism; 2) the availability of FDA-approved pharmaceuticals that target individual bioaminergic transporters and receptors85-87; and 3) the high conservation of neurotransmitter biology between invertebrates and mammals58,70,88. Further research into how these pathways can be modulated in a targeted manner could spur development of interventions that promote healthy aging and delay the onset of age-related diseases.

Methods

Strains and Growth Conditions

C. elegans were cultured according to established protocols31. In brief, worms were grown at 20 °C on standard solid nematode growth media (NGM). Throughout their lifespan, the worms were fed E. coli OP50, except during RNA interference (RNAi) experiments, where E. coli HT115 was used to deliver double-stranded RNA. Transfers of worms were carried out using a platinum wire, unless stated otherwise. The RNAi strains employed are listed in Supplementary Table 1, while the strains used in the experiments are provided in Supplementary Table 2. Genotypes were verified through PCR, and RNAi imaging results were confirmed by sequencing and quantitative PCR (qPCR) before proceeding with the experiments.

Generating transgenic strains

Neuron-specific stabilized HIF-1 strains

We used the following promoters to drive cDNA of HIF-1 (P621A)::SL2::GFP in three serotonergic neuronal populations: ADF neurons (srh-142p:: HIF-1 (P621A)), NSM neurons (ceh-2p:: HIF-1 (P621A)), HSN neurons (ham-2p:: HIF-1 (P621A)). All plasmids were verified via restriction digest and sanger sequencing, and ApE files are available upon request. Plasmids were microinjected to hif-1(ia4) KO strain by Suny Biotech using the co-injection marker myo-2p::GFP and 2-3 transgenic lines were tested in each experiment.

Neuron-specific tph-1 rescues

The pKA805[srh-142p::TPH-1] and pKA807[ceh-2p::TPH-1] constructs were generously provided by Dr. Kaveh Ashrafi. These constructs were injected at ∼50 ng/µL) with fluorescent co-injection marker myo-2p::mNeonGreen (15 ng/µL) or sur-5p::sur-5::NLSGFP (20 ng/µL) and junk DNA (up to 100 ng/µL) into gonads of day 1 gravid adult hermaphrodites. Standard protocols were followed to isolate and obtain stable over-expression mutants89.

ser-7 rescue strains

Plasmid construction and microinjection was conducted by Suny Biotech to generate all ser-7 rescue constructs. In designing the plasmids, we used the following promoters to drive cDNA of ser-7::SL2::GFP in different neuronal populations: full rescue (ser-7p::ser-7), interneuron rescue (glr-1p::ser-7), bioaminergic neuron rescue (cat-1p::ser-7), glutamatergic neuron rescue (eat-4p::ser-7), GABAergic neuron rescue (unc-47p::ser-7), GABAergic motor neuron rescue(unc-25p::ser-7), sensory neuron rescue (osm-6p::ser-7), cholinergic motor neuron rescue(acr-2p::ser-7), cholinergic neuron rescue (unc-17p::ser-7), M3 & M4 neuron rescue (ceh-28p::ser-7), and intestinal rescue (vha-6p::ser-7). All plasmids were verified via restriction digest and sanger sequencing, and ApE files are available upon request. Plasmids were microinjected by Suny Biotech using the co-injection marker myo-2p::GFP and 2-3 transgenic lines were tested in each experiment.

RIS, RIC, and BAG neuronal ablation strains

For RIS neuronal ablation strains, we purchased donor plasmid mec-18p::caspase-3 (p12)::nz [TU#813] (Plasmid #16082) and mec-18p cz::caspase-3 (p17) [TU#814] from Addgene (Plasmid #16083), and used Gibson cloning to replace mec-18p with srsx-18p. Three constructs srsx-18p::caspase-3(p12)::nz, srsx-18p::cz::caspase-3(p17), and srsx-18p::GFP were co-injected with fluorescent co-injection marker myo-3p::GFP (20 ng/µL) into the wild-type N2 strain to generate RIS genetic ablation strains. Similarly, for RIC ablation strain, tbh-1p was constructed into TU#813 and TU#814 to replace mec-18p. Three constructs tbh-1p::caspase-3(p12)::nz, tbh-1p::cz::caspase-3(p17), and tbh-1p::GFP were co-injected with fluorescent co-injection marker myo-3p::GFP (20 ng/µL) into the wild-type N2 strain to generate RIC genetic ablation strains. For BAG ablation strain, gcy-31p was constructed into TU#813 and TU#814 to replace mec-18p. Three constructs gcy-31p::caspase-3(p12)::nz, gcy-31p::cz::caspase-3(p17), and gcy-31p::GFP were co-injected with fluorescent co-injection marker myo-3p::GFP (20 ng/µL) into the wild-type N2 strain to generate BAG genetic ablation strains. All plasmids were verified via restriction digest and sanger sequencing. ApE files available upon request. Plasmid construction and microinjection was conducted by Suny Biotech.

RIM and uv1 tdc-1 rescue strains

Plasmid construction and microinjection was conducted by Suny Biotech to generate both tdc-1 rescue constructs. In designing the plasmids, we used the ocr-4 promoter to drive cDNA of tdc-1::SL2::GFP in the uv1 neuroendocrine cells. To express cDNA of tdc-1::SL2::GFP in the RIML neuron, we used the gcy-13 promoter. All plasmids were verified via restriction digest and sanger sequencing, and ApE files are available upon request. Plasmids were microinjected by Suny Biotech using the co-injection marker myo-2p::GFP.

RNAi Knockdown

For all RNAi knockdowns, worms were exposed to the RNAi treatment for two generations to achieve optimal knockdown. All RNAi clones were sourced from the Vidal RNAi library. Each RNAi clone was sequence-verified.

Quantitative PCR

For RNAi validation experiments, RNA was isolated from day 1 adult worms that had been grown on RNAi for two generations. For measuring fmo-2 induction in the hif-1(ia4); ADF::HIF-1S strain, worms were grown on OP50, synchronized, and collected at day 1 of adulthood. RNA extraction was performed using the Direct-zol RNA Miniprep Kit (Zymo), and cDNA synthesis was carried out with the Biorad iScript cDNA Synthesis Kit. Gene expression was assessed using the SYBR Green quantitative RT-PCR 90 system (BioRad), with mRNA levels normalized to the housekeeping genes cdc-42 and Y45FD10.4. Primers for RNAi validation were designed to target the 3’ UTR of each gene to avoid amplification of the bacterially produced RNAi. Gene expression was quantified using standard curves.

Lifespan Measurements

Lifespan assays were conducted following previously described protocols31. Briefly, 10-15 gravid adults were transferred to NGM plates for a three-hour timed egg lay, after which they were removed. Once the progeny reached day 1 of adulthood, 60-80 worms were placed on NGM plates containing 33 µL of 150 mM fluorodeoxyuridine (FUdR) and 100 µL of 50 mg/mL ampicillin per 100 mL of NGM. FUdR inhibits progeny development, while ampicillin prevents bacterial contamination. NGM + FUdR + Ampicillin plates were seeded with concentrated bacteria (5x for OP50-fed lifespan assays). At least two plates per strain and condition were used for each lifespan replicate. Worms were considered dead when they failed to respond to a gentle touch with a platinum wire under a dissection microscope. Lifespan data were recorded at least three times a week until all worms were dead. To prevent escape, a barrier of 75 µL of 100 mM palmitic acid (Sigma-Aldrich) dissolved in 100% ethanol was applied along the edges of each lifespan plate. Data were analyzed using R version 4.3.1 and visualized in Adobe Illustrator 2022.

RNAi Lifespans

RNAi lifespan assays were performed similarly to standard lifespan assays, with modifications to the initial timed egg lay (TEL) procedure and food concentrations. To ensure maximal RNAi knockdown, worms were first TEL’ed for 3 hours on RNAi plates. The progeny from this TEL were left on RNAi plates to develop into gravid adults and then used for a second TEL under the same RNAi conditions. Progeny from the second-generation TEL were then used for the lifespan assay. RNAi bacteria (HT115) are ampicillin-resistant and can grow slowly on AMP-containing lifespan plates. To maintain consistent bacterial availability throughout the lifespan assay, RNAi plates were seeded with bacteria at 2x concentration, starting from an optical density (OD600) of 3.0.

Fluorescent Slide Microscopy

Fluorescent images for this study were captured using a Leica M165F fluorescent microscope, controlled by Leica Application Suite X (LASX) software. A minimum of 15 worms per condition were imaged at a magnification of at least 70x. For image quantification, individual worms were carefully separated to ensure they did not touch. Custom R code91 was then used to create a pixel mask of each worm from the brightfield image and measure the fluorescent intensity of that region in the corresponding fluorescent image. The background fluorescence was subtracted from the mean fluorescence values. Data were analyzed using R version 4.3.1 and visualized in Adobe Illustrator 2022.

Pumping Assays

Pumping was measured in day 2 gravid adults, as previously described31. The number of pumps per 30 seconds was measured in at least 10 worms under 100x magnification. Because food availability is known to change pumping rate, only worms on the bacterial lawn were considered in this assay. Data were plotted by R version 4.3.1 and Adobe Illustrator 2022.

Thrashing Assays

Thrashing was measured in day 2 gravid adults, as previously described31. Briefly, approximately 10 worms were placed in a droplet of M9 solution. Once worms began moving at a maximum rate, their body bends were counted for 30 seconds. Data were plotted by R version 4.3.1 and Adobe Illustrator 2022.

30-minute Video Recordings for Velocity and Movement Patterns

On day 1 and day 5 of adulthood, 10-15 worms of each genotype were placed in the center of a 35 mm petri dish that was completely covered in 350 μL of OP50 bacteria (bacteria was grown overnight to an optical density (OD600 3.0). Worms were placed above the bright light source at the base of the microscope used for video recording for 10 minutes to acclimate to the light, and then 30-minute video recordings were started. Videos were taken at 3.65x magnification, and at 1 frame per second on a Leica M165F microscope using Leica Application Suite X (LASX) software and 4x4 binning. The videos were analyzed using the wrMTrck multiple object tracker plugin (wrMTrck_Batch v1.04) on FIJI92 (FIJI Is Just ImageJ bundled with 64-bit Java 1.8.0) to measure average speed, maximum speed, worm area, total distance traveled, and distance from the point of origin. Not all worm paths lasted the full 30-minute duration of the video due to fleeing and collisions. In cases when worms fled the plate and were no longer visible during the video, the data from that worm was included if the on-plate duration was ≥10 minutes. In cases where worms collided and wrMTrck failed to correctly determine which worm was which after the event, the data for those worms before the collision was included if the duration was ≥ 10 minutes. In all data analysis using wrMTrck data, we confirmed that there was no significant difference in average path duration per condition. Data were plotted by R version 4.3.1 and Adobe Illustrator 2022.

Statistical Analysis

The height of all bar plots represents the mean value for each condition, with error bars indicating the standard error of the mean (SEM). For comparisons involving more than two conditions, one-way ANOVA followed by post-hoc Tukey HSD tests 47 were used to assess interactions between variables and determine statistical differences between conditions. Significance levels are indicated as *p < 0.05, **p < 0.01, ***p < 0.001, and ****p < 0.0001. For lifespan assays, statistical comparisons of survivorship curves were performed using the survfit survival analysis function in R. The log-rank test was used to compare two survival curves. To assess interactions across multiple experimental variables and for comparisons involving more than two survival curves, a Cox regression analysis was conducted using the survivalMPL package in R. For experiments with multiple biological replicates, p-values were adjusted using the Bonferroni correction. Exact sample sizes (Ns) for each experiment are provided in the source data files, and minimum sample sizes for each plot are specified in the figure legends.